---
title: "roi_assess"
author: "Jackie Beltrán"
date: "2024-02-18"
output: html_document
editor_options: 
  chunk_output_type: console
---
# libraries
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(ggplot2)
require(tidyr)
require(dplyr)
require(ggpubr)
require(data.table)
require(tidyr)
require(patchwork)
require(ppcor)
```

# Data
```{r load data}
getwd()

roi <- read.csv('../../../../RL/data/roi_N74.csv') # roi extraction for each subj

assess <- read.csv('../../../../RL/data/K01_tracking_covs.csv') 

```

# preproc

```{r preproc assess}

ca_sub <- assess %>% dplyr::select('K01_SUBID', 'TEPS.Anticipatory', 'TEPS.Consummatory',
                                   'qids_score','RRS_Depression.Related',
                                   'BAS.Reward.Responsiveness..K.only.', 'BIS_score',
                                   'shaps_score..new.')

# rename column to then merge datasets
names(ca_sub)[names(ca_sub) == "K01_SUBID"] <- "subj_ID"

df_full <- merge(roi, ca_sub)
```

```{r covariates}
covs <- assess %>% dplyr::select('K01_SUBID', 'SUB_gender', 'SUB_Age', 'Medication')

# rename column
names(covs)[names(covs) == "K01_SUBID"] <- "subj_ID"

# merge
df_covs <- merge(df_full, covs)
length(unique(df_covs$subj_ID)) # N = 74

df_covs$subj_ID <- as.factor(df_covs$subj_ID)
df_covs$group<- as.factor(df_covs$group)
df_covs$SUB_gender <- as.factor(df_covs$SUB_gender)

# recode variables
df_covs <- df_covs %>% mutate(SUB_gender=recode(SUB_gender, 'F'='1', 'M'='0')) # dummy code sex
df_covs <- df_covs %>% mutate(Medication=recode(Medication, '0'='1', '1'='2')) # 1 = none, 2 = y
df_covs$SUB_gender <- as.numeric(df_covs$SUB_gender)
df_covs$Medication <- as.numeric(df_covs$Medication)
df_covs2 <- na.omit(df_covs) # remove nan

```

```{r roi}
roi_m <-
  reshape2::melt(roi,
                 id = c('subj_ID', 'group', 'ROI'),
                 variable.name = "contrast")

roi_out_3 <- roi_m %>%
  group_by(group, contrast, ROI) %>%
  mutate(
    mean_value = mean(value),
    sd_value = sd(value),
    is_outlier = value > (mean_value + 3 * sd_value) |
      value < (mean_value - 3 * sd_value)
  )

# Identify outliers 
outliers_3 <- roi_out_3 %>% filter(is_outlier == T)

# View the outliers
print(outliers_3)

roi_o <-
  anti_join(roi_out_3,
            outliers_3,
            by = c("subj_ID", "group", "contrast", "ROI", "value")) %>%
  dplyr::select("subj_ID", "group", "ROI", "contrast", "value")

roi_o_w <- reshape2::dcast(roi_o, subj_ID + group + ROI ~ contrast)
roi_o_w <- roi_o_w[order(roi_o_w$ROI, roi_o_w$subj_ID),]

col_order <-
  c(
    "group",
    "subj_ID",
    "stimGain",
    "stimLoss",
    "stimNeutral",
    "keyGo",
    "fb_loss_rpe",
    "fb_gain_rpe",
    "fb_all_rpe",
    "ROI"
  )

roi_o_w <- roi_o_w[, col_order]

```

```{r outlier free roi + regular assessments}

df_o <- merge(roi_o_w, ca_sub)

```


# VTA
```{r}
vta_ca <- df_full[df_full$ROI == "vta_mni", ]
```

```{r}
ggplot(vta_ca, aes(x=qids_score, y=fb_all_rpe, color=group)) +
  geom_point(size = 5) + 
  geom_smooth(method=lm, aes(fill=group), se = FALSE) + 
  scale_color_manual(values=c('lightblue3','burlywood2')) + 
  labs(x='Qids score', y='VTA RPE', title='') +
  theme_light() +
  theme(axis.text = element_text(size = 16)) + 
  theme(legend.position=c(0.1,0.9), legend.justification=c(0.1,0.9)) 

```

## qids
```{r}
mdd <- vta_ca[vta_ca$group == "MDD", ] # N=38
hc <- vta_ca[vta_ca$group == "HC", ] # N=36

shapiro.test(mdd$fb_all_rpe) # p = 0.2738
shapiro.test(mdd$qids_score) # p = 0.1872

ggqqplot(mdd$fb_all_rpe, ylab = "fb all rpe") # some drift near the tails but okay
ggqqplot(mdd$qids_score, ylab = "qids")

cor.test(mdd$fb_all_rpe, mdd$qids_score, 
                    method = "pearson") # p = 0.5032

cor.test(mdd$fb_gain_rpe, mdd$qids_score, 
                    method = "pearson") # p = 0.7691

cor.test(mdd$fb_loss_rpe, mdd$qids_score, 
                    method = "pearson") # p = 0.6248
```

## BAS - reward responsiveness

```{r}
ggplot(vta_ca, aes(x=fb_gain_rpe, y=BAS.Reward.Responsiveness..K.only., color=group)) +
  geom_point(size = 5) + 
  geom_smooth(method=lm, aes(fill=group), se = FALSE) + 
  scale_color_manual(values=c('lightblue3','burlywood2')) + 
  labs(x='VTA gain RPE', y='Reward Responsiveness', title='') +
  theme_light() +
  theme(axis.text = element_text(size = 16)) + 
  theme(legend.position=c(2.0,0.9), legend.justification=c(0.1,0.9)) 

```

```{r mdd}
shapiro.test(mdd$fb_all_rpe) # p = 0.2738
shapiro.test(mdd$BAS.Reward.Responsiveness..K.only.) # p = 0.1474

ggqqplot(mdd$fb_all_rpe, ylab = "fb all rpe") # some drift near the tails but okay
ggqqplot(mdd$BAS.Reward.Responsiveness..K.only., ylab = "qids")

cor.test(mdd$fb_all_rpe, mdd$BAS.Reward.Responsiveness..K.only., 
                    method = "pearson") # p = 0.1034

cor.test(hc$fb_gain_rpe, hc$BAS.Reward.Responsiveness..K.only., 
                    method = "pearson") # p = 0.6378

cor.test(mdd$fb_loss_rpe, mdd$BAS.Reward.Responsiveness..K.only., 
                    method = "pearson") # p = 0.02889 # SIG

```

```{r partial correlation}
pcor_test(
  df = df_covs2,
  roi = "vta_mni",
  contrast = "fb_all_rpe",
  scale = "BAS.Reward.Responsiveness..K.only.",
  group_col = "group",
  sex_col = "SUB_gender"
) # MDD: R = 0.2862177, p = 0.096, HC: rho = 0.2484045, p =0.150186

pcor_test(
  df = df_covs2,
  roi = "vta_mni",
  contrast = "fb_gain_rpe",
  scale = "BAS.Reward.Responsiveness..K.only.",
  group_col = "group",
  sex_col = "SUB_gender"
) # MDD: R = 0.0789, p = 0.653, HC: rho = 0.155, p = 0.373

pcor_test(
  df = df_covs2,
  roi = "vta_mni",
  contrast = "fb_loss_rpe",
  scale = "BAS.Reward.Responsiveness..K.only.",
  group_col = "group",
  sex_col = "SUB_gender"
) # MDD: R = 0.372, p = 0.0279, HC: -0.062, p = 0.725
```

```{r add age and meds as covar}

vta_sub <- df_covs2 %>% filter(ROI == "vta_mni" & group == "MDD") 
vta_sub <- df_covs2 %>% filter(ROI == "vta_mni" & group == "HC") 

pcor.test(
  vta_sub$BAS.Reward.Responsiveness..K.only.,
  vta_sub$fb_all_rpe,
  list(vta_sub$SUB_gender, vta_sub$SUB_Age, vta_sub$Medication)
  ) # R = 0.287, p = 0.106

pcor.test(
  vta_sub$BAS.Reward.Responsiveness..K.only.,
  vta_sub$fb_gain_rpe,
  list(vta_sub$SUB_gender, vta_sub$SUB_Age, vta_sub$Medication)
  ) # R = -0.045, p = 0.804

pcor.test(
  vta_sub$BAS.Reward.Responsiveness..K.only.,
  vta_sub$fb_loss_rpe,
  list(vta_sub$SUB_gender, vta_sub$SUB_Age, vta_sub$Medication),
  ) # R = 0.351, p = 0.0451
```

```{r hcc}
shapiro.test(hc$fb_all_rpe) # p = 0.03308
shapiro.test(hc$BAS.Reward.Responsiveness..K.only.) # p = 0.02521

ggqqplot(hc$fb_all_rpe, ylab = "fb all rpe") # some drift near the tails but okay
ggqqplot(hc$BAS.Reward.Responsiveness..K.only., ylab = "bas")

cor.test(hc$fb_all_rpe, hc$BAS.Reward.Responsiveness..K.only., 
                    method = "spearman") # p ~ 0.2721 (ties)

cor.test(hc$fb_gain_rpe, hc$BAS.Reward.Responsiveness..K.only., 
                    method = "spearman") # p ~ 0.5632 (ties)

cor.test(hc$fb_loss_rpe, hc$BAS.Reward.Responsiveness..K.only., 
                    method = "spearman") # p = 0.6985 

```

## BAS - BIS Score

```{r}
ggplot(mdd, aes(x=fb_all_rpe, y=BIS_score, color=group)) +
  geom_point(size = 5) + 
  geom_smooth(method=lm, aes(fill=group), se = FALSE) + 
  scale_color_manual(values=c('lightblue3','burlywood2')) + 
  labs(x='VTA RPE', y='BIS Score', title='') +
  theme_light() +
  theme(axis.text = element_text(size = 16)) + 
  theme(legend.position=c(2.0,0.9), legend.justification=c(0.1,0.9)) 

ggplot(mdd, aes(x=fb_gain_rpe, y=BIS_score, color=group)) +
  geom_point(size = 5) + 
  geom_smooth(method=lm, aes(fill=group), se = FALSE) + 
  scale_color_manual(values=c('lightblue3','burlywood2')) + 
  labs(x='VTA gain RPE', y='BIS Score', title='') +
  theme_light() +
  theme(axis.text = element_text(size = 16)) + 
  theme(legend.position=c(2.0,0.9), legend.justification=c(0.1,0.9)) # this looks maybe sig.


ggplot(mdd, aes(x=fb_loss_rpe, y=BIS_score, color=group)) +
  geom_point(size = 5) + 
  geom_smooth(method=lm, aes(fill=group), se = FALSE) + 
  scale_color_manual(values=c('lightblue3','burlywood2')) + 
  labs(x='VTA loss RPE', y='BIS Score', title='') +
  theme_light() +
  theme(axis.text = element_text(size = 16)) + 
  theme(legend.position=c(2.0,0.9), legend.justification=c(0.1,0.9)) 

```

## RRS depression related 

```{r}
shapiro.test(mdd$fb_all_rpe) # p = 0.2738
shapiro.test(mdd$RRS_Depression.Related) # p = 0.1658

ggqqplot(mdd$fb_all_rpe, ylab = "fb all rpe") # some drift near the tails but okay
ggqqplot(mdd$RRS_Depression.Related, ylab = "RRS")

cor.test(mdd$fb_all_rpe, mdd$RRS_Depression.Related, 
                    method = "pearson") # p = 0.9107

cor.test(mdd$fb_gain_rpe, mdd$RRS_Depression.Related, 
                    method = "pearson") # p = 0.5281

cor.test(mdd$fb_loss_rpe, mdd$RRS_Depression.Related, 
                    method = "pearson") # p = 0.5145 

```

## shaps
The items cover the domains of: social interaction, food and drink, sensory experience, and interest/pastimes. A score of 2 or less constitutes a “normal” score, while an “abnormal” score is defined as 3 or more. Each item has four possible responses: strongly disagree, disagree, agree, or strongly agree. 
- Someone with a high SHAPS score would also have low hedonic tone, or reduced ability to feel pleasure (https://www.verywellhealth.com/snaith-hamilton-pleasure-scale-5272382)
- 1 point = disagrees/strongly disagrees with a statement. 0 points = agrees or strongly agree
- Lower SHAPS scores indicate higher levels of hedonic capacity or a lower level of anhedonia

mdd: 
- correlation with loss rpe
- trending with full rpe such that greater signal = higher shaps score

```{r mdd}
ggplot(mdd, aes(x=fb_all_rpe, y=shaps_score..new., color=group)) +
  geom_point(size = 5) + 
  geom_smooth(method=lm, aes(fill=group), se = FALSE) + 
  scale_color_manual(values=c('lightblue3','burlywood2')) + 
  labs(x='VTA RPE', y='shaps', title='') +
  theme_light() +
  theme(axis.text = element_text(size = 16)) + 
  theme(legend.position=c(2.0,0.9), legend.justification=c(0.1,0.9)) 

shapiro.test(mdd$fb_all_rpe) # p = 0.2738
shapiro.test(mdd$fb_gain_rpe) # p = 0.7654
shapiro.test(mdd$fb_loss_rpe) # p = 0.6714
shapiro.test(mdd$shaps_score..new.) # p = 0.7385

ggqqplot(mdd$fb_all_rpe, ylab = "fb all rpe") # some drift near the tails but okay
ggqqplot(mdd$shaps_score..new., ylab = "shaps")

cor.test(mdd$fb_all_rpe, mdd$shaps_score..new., 
                    method = "pearson") # p = 0.06486, R = 0.3025374 -- TRENDING

cor.test(mdd$fb_gain_rpe, mdd$shaps_score..new., 
                    method = "pearson") # p = 0.9482

cor.test(mdd$fb_loss_rpe, mdd$shaps_score..new., 
                    method = "pearson") # p = 0.04756, R = 0.32  -- SIG


```

## outliers
```{r check for outliers}

shaps <-
  df_full %>% dplyr::select('subj_ID', 'group', 'shaps_score..new.') %>%
  distinct() # two MDD subjects missing a score

# mean HC = 18.1(4.67), mean MDD = 32.1(8.36)

shaps_m <-
  reshape2::melt(shaps,
                 id = c('subj_ID', 'group'),
                 variable.name = "shaps")

shaps_m %>% 
  group_by(group, shaps) %>%
  summarise(mean_score = mean(value), sd_score= sd(value))


shaps_out_3 <- shaps_m %>%
  group_by(group, shaps) %>%
  mutate(
    mean_value = mean(value),
    sd_value = sd(value),
    is_outlier = value > (mean_value + 3 * sd_value) |
      value < (mean_value - 3 * sd_value)
  )

# Identify outliers 
shaps_outliers <- shaps_out_3 %>% filter(is_outlier == T) # no outliers!!

```

```{r repeat analysis without roi outliers}

vta_o <- df_o[df_o$ROI == "vta_mni", ] # outlier during loss rpe 
mdd <- vta_o[vta_o$group == "MDD", ]

shapiro.test(mdd$fb_loss_rpe) # p = 0.6714
shapiro.test(mdd$shaps_score..new.) # p = 0.7385

ggqqplot(mdd$fb_loss_rpe, ylab = "fb loss rpe") # some drift near the tails but okay
ggqqplot(mdd$shaps_score..new., ylab = "shaps")

cor.test(mdd$fb_loss_rpe, mdd$shaps_score..new., 
                    method = "pearson") # p = 0.04756, R = 0.3235128 -- SIG with outlier removal


```


# BAS outlier detection 
- we remove outliers and end up with a dataframe called BAS_df
```{r}
BAS <-
  df_full %>% dplyr::select('subj_ID', 'group', 'BAS.Reward.Responsiveness..K.only.') %>%
  distinct() # two MDD subjects missing a score

BAS_2 <- na.omit(BAS)

# mean HC = 7.92(1.65), mean MDD = 10.3(3.00)

bas_m <-
  reshape2::melt(BAS_2,
                 id = c('subj_ID', 'group'),
                 variable.name = "BAS")

bas_m %>% 
  group_by(group, BAS) %>%
  summarise(mean_score = mean(value), sd_score= sd(value))

```

```{r boxplot}
ggplot(bas_m, aes(x = BAS, y = value, fill = group)) +
  geom_boxplot(outlier.shape = 19) +
    geom_dotplot(binaxis='y', stackdir='center',
                 position=position_dodge(0.9), dotsize = 0.3) +
  
  scale_fill_manual(values=c('lightblue3','peachpuff')) +

  #facet_wrap(~ROI, scales = "free_y") +
  labs(title = "",
       x = "",
       y = "score") +
  theme_minimal() + 
    theme(
    # axis.text = element_text(size = 15),
    axis.title = element_text(size = 15),
    strip.text = element_text(size = 12),
    legend.title = element_blank(),
    legend.text = element_text(size=15),
  )
```

```{r outlier removal}

bas_out_3 <- bas_m %>%
  group_by(group, BAS) %>%
  mutate(
    mean_value = mean(value),
    sd_value = sd(value),
    is_outlier = value > (mean_value + 3 * sd_value) |
      value < (mean_value - 3 * sd_value)
  )

# Identify outliers 
bas_outliers <- bas_out_3 %>% filter(is_outlier == T) # no outliers!!
```

# NAc
```{r}
nac_ca <- df_full[df_full$ROI == "rl_accumbens", ]
mdd <- nac_ca[nac_ca$group == "MDD", ] # N=38
hc <- nac_ca[nac_ca$group == "HC", ] # N=36

```

## BAS - reward responsiveness

In HC, greater RPE signal during loss trials is trending in correlation with greater reward responsiveness 
```{r}
ggplot(nac_ca, aes(x=fb_loss_rpe, y=BAS.Reward.Responsiveness..K.only., color=group)) +
  geom_point(size = 5) + 
  geom_smooth(method=lm, aes(fill=group), se = FALSE) + 
  scale_color_manual(values=c('lightblue3','burlywood2')) + 
  labs(x='NAC Loss RPE', y='BAS score', title='') +
  theme_light() +
  theme(axis.text = element_text(size = 16)) + 
  theme(legend.position=c(0.1,0.9), legend.justification=c(0.1,0.9)) 

```

```{r mdd}

shapiro.test(mdd$fb_all_rpe) # p = 0.8932
shapiro.test(mdd$fb_loss_rpe) # p = 0.3136
shapiro.test(mdd$fb_gain_rpe) # p = 0.5061
shapiro.test(mdd$BAS.Reward.Responsiveness..K.only.) # p = 0.1474

ggqqplot(mdd$fb_all_rpe, ylab = "fb all rpe")
ggqqplot(mdd$fb_loss_rpe, ylab = "fb loss rpe") # good 
ggqqplot(mdd$fb_gain_rpe, ylab = "fb gain rpe") # good
ggqqplot(mdd$BAS.Reward.Responsiveness..K.only., ylab = "BAS")

cor.test(mdd$fb_all_rpe, mdd$BAS.Reward.Responsiveness..K.only., 
                    method = "pearson") # p = 0.3839

cor.test(mdd$fb_gain_rpe, mdd$BAS.Reward.Responsiveness..K.only., 
                    method = "pearson") # p = 0.3307

cor.test(mdd$fb_loss_rpe, mdd$BAS.Reward.Responsiveness..K.only., 
                    method = "pearson") # p = 0.6843

```

```{r hc}
shapiro.test(hc$fb_all_rpe) # p = 0.004828
shapiro.test(hc$fb_loss_rpe) # p = 0.02489
shapiro.test(hc$fb_gain_rpe) # p = 0.9262
shapiro.test(hc$BAS.Reward.Responsiveness..K.only.) # p = 0.02521

ggqqplot(hc$fb_all_rpe, ylab = "fb all rpe") # drift at the tails
ggqqplot(hc$fb_loss_rpe, ylab = "fb loss rpe") # drift at the tails 
ggqqplot(hc$fb_gain_rpe, ylab = "fb gain rpe") # good
ggqqplot(hc$BAS.Reward.Responsiveness..K.only., ylab = "BAS")

cor.test(hc$fb_all_rpe, hc$BAS.Reward.Responsiveness..K.only., 
                    method = "spearman") # p = 0.372

cor.test(hc$fb_gain_rpe, hc$BAS.Reward.Responsiveness..K.only., 
                    method = "spearman") # p = 0.6676

cor.test(hc$fb_loss_rpe, hc$BAS.Reward.Responsiveness..K.only., 
                    method = "spearman") # p = 0.1676

```

## shaps

```{r}
ggplot(mdd, aes(x=fb_all_rpe, y=shaps_score..new., color=group)) +
  geom_point(size = 5) + 
  geom_smooth(method=lm, aes(fill=group), se = FALSE) + 
  scale_color_manual(values=c('lightblue3','burlywood2')) + 
  labs(x='NAC RPE', y='shaps', title='') +
  theme_light() +
  theme(axis.text = element_text(size = 16)) + 
  theme(legend.position=c(0.1,0.9), legend.justification=c(0.1,0.9)) 

shapiro.test(mdd$fb_all_rpe) # p = 0.8932
shapiro.test(mdd$fb_loss_rpe) # p = 0.3136
shapiro.test(mdd$fb_gain_rpe) # p = 0.5061
shapiro.test(mdd$shaps_score..new.) # p = .7385

ggqqplot(mdd$fb_all_rpe, ylab = "fb all rpe")
ggqqplot(mdd$fb_loss_rpe, ylab = "fb loss rpe") # good 
ggqqplot(mdd$fb_gain_rpe, ylab = "fb gain rpe") # good
ggqqplot(mdd$shaps_score..new., ylab = "shaps") # good

cor.test(mdd$fb_all_rpe, mdd$shaps_score..new., 
                    method = "pearson") # p = 0.5953

cor.test(mdd$fb_gain_rpe, mdd$shaps_score..new., 
                    method = "pearson") # p = 0.383

cor.test(mdd$fb_loss_rpe, mdd$shaps_score..new., 
                    method = "pearson") # p = 0.9531


```

