---
title: "roi_firstlevel_rpe"
author: "Jackie Beltrán"
date: "2023-02-02"
output: html_document
editor_options: 
  chunk_output_type: console
---

# libraries
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(ggplot2)
require(tidyr)
require(dplyr)
require(ggpubr)
require(data.table)
require(tidyr)
require(patchwork)
require(ppcor)
```

# Data
```{r load data}
getwd()

roi <- read.csv('../../../../RL/data/roi_N74.csv') # roi extraction for each subj

assess <- read.csv('../../../../RL/data/K01_tracking_covs.csv') 

```

# Functions 

```{r function}

# correlation with clinical assessments 
ca_analysis <- function(df, roi, contrast, scale, group_col) {
  
  # subset the dataframe for the ROI and contrast of interest
  
  df_sub = df[df$ROI == roi,c(group_col, contrast, scale)]
  
  HC_normalcy_contrast_test <-
    with(df_sub, shapiro.test(df_sub[[contrast]][group == "HC"]))
  
  MDD_normalcy_contrast_test <-
    with(df_sub, shapiro.test(df_sub[[contrast]][group == "MDD"]))
  
  HC_normalcy_scale_test <- with(df_sub, shapiro.test(df_sub[[scale]][group == "HC"]))
  MDD_normalcy_scale_test <- with(df_sub, shapiro.test(df_sub[[scale]][group == "MDD"]))

  if (HC_normalcy_contrast_test$p.value < 0.05 |
      HC_normalcy_scale_test$p.value < 0.05) {
    print("Conducting non-parametric test...")
    hc_spearman_result <-
      cor.test(df_sub[[contrast]][df_sub[[group_col]] == "HC"],
               df_sub[[scale]][df_sub[[group_col]] == "HC"], method = "spearman")
    hc_pearson_result <- NULL
    
  } else {
    print("Conducting pearson correlation test...")
    hc_spearman_result <- NULL
    hc_pearson_result <-
      cor.test(df_sub[[contrast]][df_sub[[group_col]] == "HC"],
               df_sub[[scale]][df_sub[[group_col]] == "HC"], 
               method = "pearson")
  }
  
  if (MDD_normalcy_contrast_test$p.value < 0.05 |
    MDD_normalcy_scale_test$p.value < 0.05) {
    print("Conducting non-parametric test...")
    
    mdd_spearman_result <-
      cor.test(df_sub[[contrast]][df_sub[[group_col]] == "MDD"],
               df_sub[[scale]][df_sub[[group_col]] == "MDD"], 
               method = "spearman")
    mdd_pearson_result <- NULL
  } else {
    print("Conducting pearson correlation test...")
    mdd_spearman_result <- NULL
    mdd_pearson_result <-
      cor.test(df_sub[[contrast]][df_sub[[group_col]] == "MDD"],
               df_sub[[scale]][df_sub[[group_col]] == "MDD"], 
               method = "pearson")
  }

  results <- list(
    HC_normalcy_contrast_test = HC_normalcy_contrast_test,
    HC_normalcy_scale_test = HC_normalcy_scale_test,
    MDD_normalcy_contrast_test = MDD_normalcy_contrast_test,
    MDD_normalcy_scale_test = MDD_normalcy_scale_test,
    hc_spearman_result = hc_spearman_result,
    hc_pearson_result = hc_pearson_result,
    mdd_spearman_result = mdd_spearman_result,
    mdd_pearson_result = mdd_pearson_result
  )

  return(results)
}

```

# preproc

```{r preproc assess}

ca_sub <- assess %>% dplyr::select('K01_SUBID', 'TEPS.Anticipatory', 'TEPS.Consummatory')

# rename column to then merge datasets
names(ca_sub)[names(ca_sub) == "K01_SUBID"] <- "subj_ID"

df_full <- merge(roi, ca_sub)
```

## TEPS
- we remove outliers and end up with a dataframe called teps_df
```{r}
teps <-
  df_full %>% dplyr::select('subj_ID', 'group', 'TEPS.Anticipatory', 'TEPS.Consummatory') %>% 
  distinct()

# anticipatory; mean HC = 45.9(8.75), mean MDD = 35.7(10.2)
# consum; mean HC = 38.8(7.93), mean MDD = 32.9(9.16)
teps_m <-
  reshape2::melt(teps,
                 id = c('subj_ID', 'group'),
                 variable.name = "TEPS")

teps_sum <- teps_m %>% 
  group_by(group, TEPS) %>%
  summarise(mean_contrast = mean(value), sd_contrast= sd(value))

```

```{r boxplot}
ggplot(teps_m, aes(x = TEPS, y = value, fill = group)) +
  geom_boxplot(outlier.shape = NA) +
    geom_dotplot(binaxis='y', stackdir='center',
                 position=position_dodge(0.9), dotsize = 0.3) +
  
  scale_fill_manual(values=c('lightblue3','peachpuff')) +

  #facet_wrap(~ROI, scales = "free_y") +
  labs(title = "",
       x = "",
       y = "score") +
  theme_minimal() + 
    theme(
    # axis.text = element_text(size = 15),
    axis.title = element_text(size = 15),
    strip.text = element_text(size = 12),
    legend.title = element_blank(),
    legend.text = element_text(size=15),
  )
```

```{r outlier removal}

teps_out_3 <- teps_m %>%
  group_by(group, TEPS) %>%
  mutate(
    mean_value = mean(value),
    sd_value = sd(value),
    is_outlier = value > (mean_value + 3 * sd_value) |
      value < (mean_value - 3 * sd_value)
  )

# Identify outliers 
teps_outliers <- teps_out_3 %>% filter(is_outlier == T)

# View the outliers
print(teps_outliers) # HC, 543 on both scales: teps_ant = 15, teps_cons = 9

# remove outliers
teps_df <-
  anti_join(teps_out_3,
            teps_outliers,
            by = c("subj_ID", "group", "TEPS")) %>%
  dplyr::select("subj_ID", "group", "TEPS", "value")

# quick check, subj 543 should be omitted in teps_df
teps_df %>% filter(subj_ID == 543)

## summary statistics for dataset without outliers
teps_df_sum <- teps_df %>% 
  group_by(group, TEPS) %>%
  summarise(mean_value = mean(value), sd_value= sd(value))

```

## ROI
- we remove outliers and end up with df called roi_o
```{r subset data}
# select for ROIs of interest
roi_filtered <- roi %>%
  filter(
    ROI %in% c(
      "vta_mni",
      "vta_aveHCPT",
      "amyg",
      "LB_amyg",
      "vmPFC",
      "rl_accumbens",
      "antrInsula"
    )
  )

length(unique(roi_filtered$subj_ID)) # N = 74

str(roi_filtered)
roi_filtered$subj_ID <- as.factor(roi_filtered$subj_ID)
roi_filtered$group <- as.factor(roi_filtered$group)
roi_filtered$ROI <- as.factor(roi_filtered$ROI)

# subset for MDD
roi_mdd <- roi_filtered %>% filter(grepl("^4", subj_ID))
length(unique(roi_mdd$subj_ID)) # N = 38

# melt just to plot / generate a summary of means & stds
roi_contrasts_m <-
  reshape2::melt(roi_filtered,
                 id = c('subj_ID', 'group', 'ROI'),
                 variable.name = "contrast")

roi_contrasts_sum <- roi_contrasts_m %>% 
  group_by(group, contrast, ROI) %>%
  summarise(mean_contrast = mean(value), sd_contrast= sd(value)) %>% arrange(contrast, ROI)

# write.csv(roi_contrasts_sum, file = "roi_summary_stats.csv", row.names=F)

```

```{r box plot}

ggplot(roi_contrasts_m, aes(x = contrast, y = value, fill = group)) +
  geom_boxplot(outlier.shape = NA) +
    geom_dotplot(binaxis='y', stackdir='center',
                 position=position_dodge(0.9), dotsize = 0.3) +
  
  scale_fill_manual(values=c('lightblue3','peachpuff')) +

  facet_wrap(~ROI, scales = "free_y") +
  labs(title = "ROI signal intensity by contrast and group",
       x = "Contrast",
       y = "Signal") +
  theme_minimal() + 
    theme(
    # axis.text = element_text(size = 15),
    axis.title = element_text(size = 15),
    strip.text = element_text(size = 12),
    legend.title = element_blank(),
    legend.text = element_text(size=15),
  )

```

```{r plot mean + sd}
ggplot(roi_contrasts_sum, aes(x = contrast, y = mean_contrast, fill = group)) +
  geom_bar(stat = "identity", position = "dodge") +
  
  geom_errorbar(
    aes(ymin = mean_contrast - sd_contrast, ymax = mean_contrast + sd_contrast),
    width = .1,
    position = position_dodge(.9)
  ) +
  
  scale_fill_manual(values=c('lightblue3','peachpuff')) +

  facet_wrap(~ROI, scales = "free_y") +
  labs(title = "Mean ROI signal intensity by contrast and group",
       x = "Contrast",
       y = "Signal") +
  theme_minimal() + 
    theme(
    axis.text.x = element_text(angle=45, hjust=1),
    axis.title = element_text(size = 15),
    strip.text = element_text(size = 12),
    legend.title = element_blank(),
    legend.text = element_text(size=15),
  )

```


```{r detect outliers}

roi_out_3 <- roi_contrasts_m %>%
  group_by(group, contrast, ROI) %>%
  mutate(
    mean_value = mean(value),
    sd_value = sd(value),
    is_outlier = value > (mean_value + 3 * sd_value) |
      value < (mean_value - 3 * sd_value)
  )

# Identify outliers 
outliers_3 <- roi_out_3 %>% filter(is_outlier == T)

# View the outliers
print(outliers_3)
```

```{r remove outliers & generate summary}
roi_o <-
  anti_join(roi_out_3,
            outliers_3,
            by = c("subj_ID", "group", "contrast", "ROI", "value")) %>%
  dplyr::select("subj_ID", "group", "ROI", "contrast", "value")

# quick check, subj 413 nac, stimgain should be omitted in roi_o
roi_o %>% filter(subj_ID == 413 & ROI == "rl_accumbens")

```

```{r summary }

## summary statistics for dataset without outliers
roi_o_sum <- roi_o %>% 
  group_by(group, contrast, ROI) %>%
  summarise(mean_contrast = mean(value), sd_contrast= sd(value)) %>% arrange(contrast, ROI)

```


```{r reshape and merge}
roi_o_w <- reshape2::dcast(roi_o, subj_ID + group + ROI ~ contrast)
roi_o_w <- roi_o_w[order(roi_o_w$ROI, roi_o_w$subj_ID),]

col_order <-
  c(
    "group",
    "subj_ID",
    "stimGain",
    "stimLoss",
    "stimNeutral",
    "keyGo",
    "fb_loss_rpe",
    "fb_gain_rpe",
    "fb_all_rpe",
    "ROI"
  )

roi_o_w <- roi_o_w[, col_order]

teps_o_w <- reshape2::dcast(teps_df, subj_ID + group ~ TEPS)

```

- generate a df free of outliers from the TEPS and ROI dataframe
```{r merge}
df <- merge(teps_o_w, roi_o_w)

length(unique(df$subj_ID)) # N = 73, one subject missing

str(df)

```

- generate df free of outliers that includes covariates, df_covs
```{r add covariates}

covs <- assess %>% dplyr::select('K01_SUBID', 'SUB_gender', 'SUB_Age', 'Medication')

# rename column
names(covs)[names(covs) == "K01_SUBID"] <- "subj_ID"

# merge
df_covs <- merge(df, covs)
length(unique(df_covs$subj_ID)) # N = 73


df_covs$subj_ID <- as.factor(df_covs$subj_ID)
df_covs$group<- as.factor(df_covs$group)
df_covs$SUB_gender <- as.factor(df_covs$SUB_gender)

```

```{r subset subjects without medication}

df_covs_nomed <- df_covs %>% filter(Medication == "1")

```

# ASSESSMENT CORRELATIONS
- use df
## VTA
- Consummatory anhedonia

HC: R=-0.283, p = 0.161
MDD: R = -0.13, p = 0.53

```{r stats}

# Consummatory anhedonia & vta 

vta_ca <- df[df$ROI == "vta_mni", ]

hc <- vta_ca[vta_ca$group == "HC", ] # N=35
mdd <- vta_ca[vta_ca$group == "MDD", ] # N=38

# the data are approx normally distributed

shapiro.test(hc$fb_gain_rpe) # => p = 0.14
shapiro.test(hc$fb_loss_rpe) # p = 0.78
shapiro.test(hc$fb_all_rpe) # p = 0.04 # not normally distributed
shapiro.test(hc$TEPS.Consummatory) # => p = 0.052

shapiro.test(mdd$fb_gain_rpe) # p = 0.77
shapiro.test(mdd$fb_loss_rpe) # p = 0.67
shapiro.test(mdd$fb_all_rpe) # p = 0.274
shapiro.test(mdd$TEPS.Consummatory) # p = 0.45

ggqqplot(hc$fb_gain_rpe, ylab = "fb gain rpe") # line is completely flat
ggqqplot(hc$fb_loss_rpe, ylab = "fb loss rpe") # line is completely flat
ggqqplot(hc$fb_all_rpe, ylab = "fb all rpe") # some drift near the tails
ggqqplot(hc$TEPS.Consummatory, ylab = "hc teps cons") # potential outliers

ggqqplot(mdd$fb_gain_rpe, ylab = "fb gain rpe") # looks good
ggqqplot(mdd$fb_loss_rpe, ylab = "fb loss rpe") # looks good
ggqqplot(mdd$fb_all_rpe, ylab = "fb all rpe") # some drift near the tails but okay
ggqqplot(mdd$TEPS.Consummatory, ylab = "mdd teps cons") # looks good

# HC
res <- cor.test(hc$fb_gain_rpe, hc$TEPS.Consummatory, 
                    method = "pearson")
res
res$p.value # 0.58

# MDD
res <- cor.test(mdd$fb_gain_rpe, mdd$TEPS.Consummatory, 
                    method = "pearson")
res
res$p.value # 0.6777

```

```{r loss rpe }
# HC
res <- cor.test(hc$fb_loss_rpe, hc$TEPS.Consummatory, 
                    method = "pearson")
res
res$p.value # 0.09

# MDD
res <- cor.test(mdd$fb_loss_rpe, mdd$TEPS.Consummatory, 
                    method = "pearson")
res # p = 0.3521

```


```{r full rpe}

# HC
res <- cor.test(hc$fb_all_rpe, hc$TEPS.Consummatory, 
                    method = "spearman")
res
res$p.value # 0.082

# MDD
res <- cor.test(mdd$fb_all_rpe, mdd$TEPS.Consummatory, 
                    method = "pearson")
res # p = 0.599

```

```{r plot}
### plot

# consummatory & gain RPE
ggplot(vta_ca, aes(x=TEPS.Consummatory, y=fb_gain_rpe, color=group)) +
  geom_point(size = 5) + 
  geom_smooth(method=lm, aes(fill=group), se = FALSE) + 
  scale_color_manual(values=c('lightblue3','burlywood2')) + 
  labs(x='Consummatory Anhedonia', y='Gain condition RPE', title='') +
  theme_light() +
  theme(axis.text = element_text(size = 16)) + 
  theme(legend.position=c(0.1,0.9), legend.justification=c(0.1,0.9)) 

# consummatory & loss RPE 
ggplot(vta_ca, aes(x=TEPS.Consummatory, y=fb_loss_rpe, color=group)) +
  geom_point(size = 5) + 
  geom_smooth(method=lm, aes(fill=group), se = FALSE) + 
  scale_color_manual(values=c('lightblue3','burlywood2')) + 
  labs(x='Consummatory Anhedonia', y='Loss condition RPE', title='') +
  theme_light() +
  theme(axis.text = element_text(size = 16)) + 
  theme(legend.position=c(0.1,0.9), legend.justification=c(0.1,0.9)) 

# consummatory & full RPE
ggplot(vta_ca, aes(x=TEPS.Consummatory, y=fb_all_rpe, color=group)) +
  geom_point(size = 5) + 
  geom_smooth(method=lm, aes(fill=group), se = FALSE) + 
  scale_color_manual(values=c('lightblue3','burlywood2')) + 
  labs(x='Consummatory Anhedonia', y='RPE', title='') +
  theme_light() +
  theme(axis.text = element_text(size = 16)) + 
  theme(legend.position=c(0.1,0.9), legend.justification=c(0.1,0.9)) 

# anticipatory & gain rpe

vta_gain <- ggplot(vta_ca, aes(x=TEPS.Anticipatory, y=fb_gain_rpe, color=group)) +
  geom_point(size = 5) + 
  geom_smooth(method=lm, aes(fill=group), se = FALSE, fullrange=TRUE) + 
  scale_color_manual(values=c('lightblue3','burlywood2')) + 
  labs(x='Anticipatory Anhedonia', y='Gain condition RPE', title='') +
  theme_light() +
  theme(axis.text = element_text(size = 16)) + 
  theme(legend.position=c(0.1,0.9), legend.justification=c(0.1,0.9)) 

# anticipatory & loss rpe
vta_loss <- ggplot(vta_ca, aes(x=TEPS.Anticipatory, y=fb_loss_rpe, color=group)) +
  geom_point(size = 5) + 
  geom_smooth(method=lm, aes(fill=group), se = FALSE, fullrange=TRUE) + 
  scale_color_manual(values=c('lightblue3','burlywood2')) + 
  labs(x='Anticipatory Anhedonia', y='Loss condition RPE', title='') +
  theme_light() +
  theme(axis.text = element_text(size = 16)) + 
  theme(legend.position=c(0.1,0.9), legend.justification=c(0.1,0.9)) 

# anticipatory & full rpe
vta_full <- ggplot(vta_ca, aes(x=TEPS.Anticipatory, y=fb_all_rpe, color=group)) +
  geom_point(size = 5) + 
  geom_smooth(method=lm, aes(fill=group), se = FALSE, fullrange=TRUE) + 
  scale_color_manual(values=c('lightblue3','burlywood2')) + 
  labs(x='Anticipatory Anhedonia', y='RPE', title='') +
  theme_light() +
  theme(axis.text = element_text(size = 16)) + 
  theme(legend.position=c(0.1,0.9), legend.justification=c(0.1,0.9)) 


```

```{r}
vta_full + vta_loss + vta_gain
```

Anticipatory anhedonia 

```{r stats, gain rpe}

# are the data are normally distributed

shapiro.test(hc$fb_gain_rpe) # => p = 0.14
shapiro.test(hc$TEPS.Anticipatory) # => p = 0.25

shapiro.test(mdd$fb_gain_rpe) # p = 0.77
shapiro.test(mdd$TEPS.Anticipatory) # p = 0.5988

# HC
res <- cor.test(hc$fb_gain_rpe, hc$TEPS.Anticipatory, 
                    method = "pearson")
res
res$p.value # 0.47

# MDD
res <- cor.test(mdd$fb_gain_rpe, mdd$TEPS.Anticipatory, 
                    method = "pearson")
res
res$p.value # 0.58

```

```{r stats, loss rpe}

# the data are normally distributed -- yes, the p values are greater than 0.05 

shapiro.test(hc$fb_loss_rpe) # p = 0.78
shapiro.test(hc$TEPS.Anticipatory) # p = 0.2493
shapiro.test(mdd$fb_loss_rpe) # p = 0.671
shapiro.test(mdd$TEPS.Anticipatory) # p = 0.599

# HC
res <- cor.test(hc$fb_loss_rpe, hc$TEPS.Anticipatory, 
                    method = "pearson")
res
res$p.value # 0.6206424

# MDD
res <- cor.test(mdd$fb_loss_rpe, mdd$TEPS.Anticipatory, 
                    method = "pearson")
res
res$p.value # 0.486

```

```{r full rpe}
shapiro.test(hc$fb_all_rpe) # p = 0.049
shapiro.test(hc$TEPS.Anticipatory) # p = 0.2493
shapiro.test(mdd$fb_all_rpe) # p = 0.27
shapiro.test(mdd$TEPS.Anticipatory) # p = 0.599

# HC
res <- cor.test(hc$fb_all_rpe, hc$TEPS.Anticipatory, 
                    method = "spearman")
res
res$p.value # 0.5627

# MDD
res <- cor.test(mdd$fb_all_rpe, mdd$TEPS.Anticipatory, 
                    method = "pearson")
res
res$p.value # 0.8339
```


## VTA_ave 

```{r consummatory}

# fb gain
# HC: t = -0.28091, df = 33, p-value = 0.7805 ; MDD: t = 0.51153, df = 36, p-value = 0.6121
ca_analysis(df = df, roi = "vta_aveHCPT", contrast = "fb_gain_rpe", 
            scale = "TEPS.Consummatory", group_col = "group")

# fb loss
# HC: t = -1.8919, df = 33, p-value = 0.06731 ; MDD: t = -0.19574, df = 35, p-value = 0.8459
ca_analysis(df = df, roi = "vta_aveHCPT", contrast = "fb_loss_rpe", 
            scale = "TEPS.Consummatory", group_col = "group")

# full rpe
# HC: t = -1.7828, df = 33, p-value = 0.08382 ; MDD: t = -0.27994, df = 36, p-value = 0.7811
ca_analysis(df = df, roi = "vta_aveHCPT", contrast = "fb_all_rpe", 
            scale = "TEPS.Consummatory", group_col = "group")


```

```{r anticipatory}
# fb gain
# HC: t = 0.3876, df = 33, p-value = 0.7008
# MDD: t = 0.14141, df = 36, p-value = 0.8883
ca_analysis(df = df, roi = "vta_aveHCPT", contrast = "fb_gain_rpe", 
            scale = "TEPS.Anticipatory", group_col = "group")

# fb loss
# HC: t = -1.2123, df = 33, p-value = 0.234
# MDD: t = -0.091195, df = 35, p-value = 0.9279
ca_analysis(df = df, roi = "vta_aveHCPT", contrast = "fb_loss_rpe", 
            scale = "TEPS.Anticipatory", group_col = "group")

# full rpe
# HC: t = -0.71771, df = 33, p-value = 0.478
# MDD: t = -0.64652, df = 36, p-value = 0.522
ca_analysis(df = df, roi = "vta_aveHCPT", contrast = "fb_all_rpe", 
            scale = "TEPS.Anticipatory", group_col = "group")

```

## NAC 

```{r}
# fb gain
# HC: t = 1.7197, df = 33, p-value = 0.09486
# MDD: t = -0.83718, df = 36, p-value = 0.408
ca_analysis(df = df, roi = "rl_accumbens", contrast = "fb_gain_rpe", 
            scale = "TEPS.Consummatory", group_col = "group")

# fb loss - ties issue
# error
ca_analysis(df = df, roi = "rl_accumbens", contrast = "fb_loss_rpe", 
            scale = "TEPS.Consummatory", group_col = "group")

nac_ca <- df[df$ROI == "rl_accumbens", ]

hc <- nac_ca[nac_ca$group == "HC", ] # N=35
mdd <- nac_ca[nac_ca$group == "MDD", ] # N=38

shapiro.test(hc$fb_loss_rpe) # p = 0.02
shapiro.test(hc$TEPS.Consummatory) # p = 0.05
cor.test(hc$TEPS.Consummatory, hc$fb_loss_rpe, method = "spearman") # S = 6208.8, p = 0.4552 

shapiro.test(mdd$fb_loss_rpe) # p = 0.3136
shapiro.test(mdd$TEPS.Consummatory) # p = 0.45
cor.test(mdd$TEPS.Consummatory, mdd$fb_loss_rpe, method = "pearson") # p = 0.2872

# full rpe - ties issue
ca_analysis(df = df, roi = "rl_accumbens", contrast = "fb_all_rpe", 
            scale = "TEPS.Consummatory", group_col = "group")

shapiro.test(hc$fb_all_rpe) # 0.006344
cor.test(hc$TEPS.Consummatory, hc$fb_all_rpe, method = "spearman") # p = 0.1508

shapiro.test(mdd$fb_all_rpe) # 0.8932
cor.test(mdd$TEPS.Consummatory, mdd$fb_all_rpe, method = "pearson") # p = 0.7229

```

## vmPFC - sig anticipatory results

```{r consummatory}

# fb gain
# HC: p = 0.2138, MDD: p = 0.2116
ca_analysis(df = df, roi = "vmPFC", contrast = "fb_gain_rpe", 
            scale = "TEPS.Consummatory", group_col = "group")

# fb loss
# HC: p = 0.6082 ; MDD, p = 0.9337
ca_analysis(df = df, roi = "vmPFC", contrast = "fb_loss_rpe", 
            scale = "TEPS.Consummatory", group_col = "group")

# full rpe
# HC: t = -1.7828, df = 33, p-value = 0.08382 ; MDD: t = -0.27994, df = 36, p-value = 0.7811
ca_analysis(df = df, roi = "vmPFC", contrast = "fb_all_rpe", 
            scale = "TEPS.Consummatory", group_col = "group")


```

```{r anticipatory}
# fb gain
# HC: p = 0.9207, MDD: p = 0.04017
ca_analysis(df = df, roi = "vmPFC", contrast = "fb_gain_rpe", 
            scale = "TEPS.Anticipatory", group_col = "group")

# fb loss
# HC: p = 0.4381 ; MDD, p = 0.6359
ca_analysis(df = df, roi = "vmPFC", contrast = "fb_loss_rpe", 
            scale = "TEPS.Anticipatory", group_col = "group")

# full rpe - warning with ties
# HC: 
ca_analysis(df = df, roi = "vmPFC", contrast = "fb_all_rpe", 
            scale = "TEPS.Anticipatory", group_col = "group")

# subset data 
vmpfc_ca <- df[df$ROI == "vmPFC", ]
hc <- vmpfc_ca[vmpfc_ca$group == "HC", ] # N=35
mdd <- vmpfc_ca[vmpfc_ca$group == "MDD", ] # N=38

shapiro.test(hc$fb_all_rpe) # 0.00368
cor.test(hc$TEPS.Anticipatory, hc$fb_all_rpe, method = "spearman") # p = 0.3681 -- with ties

shapiro.test(mdd$fb_all_rpe) # 0.2367
shapiro.test(mdd$TEPS.Anticipatory) # 0.5988
cor.test(mdd$TEPS.Anticipatory, mdd$fb_all_rpe, method = "pearson") # p = 0.1747


```

```{r plot}
ggplot(vmpfc_ca, aes(x=TEPS.Anticipatory, y=fb_gain_rpe, color=group)) +
  geom_point(size = 5) + 
  geom_smooth(method=lm, aes(fill=group), se = FALSE, fullrange=TRUE) + 
  scale_color_manual(values=c('lightblue3','burlywood2')) + 
  labs(x='Anticipatory Anhedonia', y='vmPFC Gain condition RPE', title='') +
  theme_light() +
  theme(axis.text = element_text(size = 16)) + 
  theme(legend.position=c(0.1,0.9), legend.justification=c(0.1,0.9)) 

# anticipatory & loss rpe
ggplot(vmpfc_ca, aes(x=TEPS.Anticipatory, y=fb_loss_rpe, color=group)) +
  geom_point(size = 5) + 
  geom_smooth(method=lm, aes(fill=group), se = FALSE, fullrange=TRUE) + 
  scale_color_manual(values=c('lightblue3','burlywood2')) + 
  labs(x='Anticipatory Anhedonia', y='vmPFC Loss condition RPE', title='') +
  theme_light() +
  theme(axis.text = element_text(size = 16)) + 
  theme(legend.position=c(0.1,0.9), legend.justification=c(0.1,0.9)) 

# anticipatory & full rpe
ggplot(vmpfc_ca, aes(x=TEPS.Anticipatory, y=fb_all_rpe, color=group)) +
  geom_point(size = 5) + 
  geom_smooth(method=lm, aes(fill=group), se = FALSE, fullrange = T) + 
  scale_color_manual(values=c('lightblue3','burlywood2')) + 
  labs(x='Anticipatory Anhedonia', y='vmPFC RPE', title='') +
  theme_light() +
  theme(axis.text = element_text(size = 16)) + 
  theme(legend.position=c(0.1,0.9), legend.justification=c(0.1,0.9)) 

```


## Amygdala
Consummatory
```{r gain RPE}

amyg_ca <- df[df$ROI == "amyg", ]

hc <- amyg_ca[amyg_ca$group == "HC", ]
mdd <- amyg_ca[amyg_ca$group == "MDD", ]

# HC
res <- cor.test(hc$fb_gain_rpe, hc$teps_cons, 
                    method = "pearson")
res
res$p.value # p=0.73

# MDD
res <- cor.test(mdd$fb_gain_rpe, mdd$teps_cons, 
                    method = "pearson")
res # p = 0.39

```

```{r loss RPE}
# HC
res <- cor.test(hc$fb_loss_rpe, hc$teps_cons, 
                    method = "pearson")
res # p-value = 0.2416

# MDD
res <- cor.test(mdd$fb_loss_rpe, mdd$teps_cons, 
                    method = "pearson")
res # p-value = 0.9205

```

Anticipatory 

```{r gain RPE}

# HC
res <- cor.test(hc$fb_gain_rpe, hc$teps_ant, 
                    method = "pearson")
res # p-value = 0.8626

# MDD
res <- cor.test(mdd$fb_gain_rpe, mdd$teps_ant, 
                    method = "pearson")
res # p-value = 0.4012

```

```{r loss rpe}

# HC
res <- cor.test(hc$fb_loss_rpe, hc$teps_ant, 
                    method = "pearson")
res # p-value = 0.6725

# MDD
res <- cor.test(mdd$fb_loss_rpe, mdd$teps_ant, 
                    method = "pearson")
res # p-value = 0.8678

```

## Habenula 

```{r consummatory gain rpe}
hab_ca <- df[df$ROI == "hb_mni", ]

hc <- hab_ca[hab_ca$group == "HC", ] # N=35
mdd <- hab_ca[hab_ca$group == "MDD", ] # N=37

# HC
res <- cor.test(hc$fb_gain_rpe, hc$teps_cons, 
                    method = "pearson")
res # p = 0.07

# MDD
res <- cor.test(mdd$fb_gain_rpe, mdd$teps_cons, 
                    method = "pearson")
res # p = 0.9165

```

```{r loss rpe}
# HC
res <- cor.test(hc$fb_loss_rpe, hc$teps_cons, 
                    method = "pearson")
res # p = 0.0607

# MDD
res <- cor.test(mdd$fb_loss_rpe, mdd$teps_cons, 
                    method = "pearson")
res # p = 0.373

```

- Anticipatory
```{r anticipatory gain rpe - HC sig}
# HC
res <- cor.test(hc$fb_gain_rpe, hc$teps_ant, 
                    method = "pearson")
res # p = 0.0076, R= -0.44

# MDD
res <- cor.test(mdd$fb_gain_rpe, mdd$teps_ant, 
                    method = "pearson")
res # p = 0.6889, R = -0.068

```

```{r loss rpe}
# HC
res <- cor.test(hc$fb_loss_rpe, hc$teps_ant, 
                    method = "pearson")
res # p = 0.33

# MDD
res <- cor.test(mdd$fb_loss_rpe, mdd$teps_ant, 
                    method = "pearson")
res # p = 0.44

```

```{r plots}
# consummatory & gain RPE
ggplot(hab_ca, aes(x=teps_cons, y=fb_gain_rpe, color=group)) +
  geom_point(size = 5) + 
  geom_smooth(method=lm, aes(fill=group), se = FALSE) + 
  scale_color_manual(values=c('lightblue3','burlywood2')) + 
  labs(x='Consummatory Anhedonia', y='Habenula Gain condition RPE', title='') +
  theme_light() +
  theme(axis.text = element_text(size = 16)) + 
  theme(legend.position=c(0.1,0.9), legend.justification=c(0.1,0.9)) 

# consummatory & loss RPE 
ggplot(hab_ca, aes(x=teps_cons, y=fb_loss_rpe, color=group)) +
  geom_point(size = 5) + 
  geom_smooth(method=lm, aes(fill=group), se = FALSE) + 
  scale_color_manual(values=c('lightblue3','burlywood2')) + 
  labs(x='Consummatory Anhedonia', y='Habenula Loss condition RPE', title='') +
  theme_light() +
  theme(axis.text = element_text(size = 16)) + 
  theme(legend.position=c(0.1,0.9), legend.justification=c(0.1,0.9)) 

# anticipatory & gain rpe

ggplot(hab_ca, aes(x=teps_ant, y=fb_gain_rpe, color=group)) +
  geom_point(size = 5) + 
  geom_smooth(method=lm, aes(fill=group), se = FALSE) + 
  scale_color_manual(values=c('lightblue3','burlywood2')) + 
  labs(x='Anticipatory Anhedonia', y='Habenula Gain condition RPE', title='') +
  theme_light() +
  theme(axis.text = element_text(size = 16)) + 
  theme(legend.position=c(0.1,0.9), legend.justification=c(0.1,0.9)) 

# anticipatory & loss rpe
ggplot(hab_ca, aes(x=teps_ant, y=fb_loss_rpe, color=group)) +
  geom_point(size = 5) + 
  geom_smooth(method=lm, aes(fill=group), se = FALSE) + 
  scale_color_manual(values=c('lightblue3','burlywood2')) + 
  labs(x='Consummatory Anhedonia', y='Habenula Loss condition RPE', title='') +
  theme_light() +
  theme(axis.text = element_text(size = 16)) + 
  theme(legend.position=c(0.1,0.9), legend.justification=c(0.1,0.9)) 

```

```{r}
ggplot(hc, aes(x=teps_ant, y=fb_gain_rpe)) +
  geom_point(size = 5) + 
  geom_smooth(method=lm, aes(fill=group), se = FALSE) + 
  scale_color_manual(values=c('lightblue3','burlywood2')) + 
  labs(x='Anticipatory Anhedonia', y='Habenula Gain condition RPE', title='') +
  theme_light() +
  theme(axis.text = element_text(size = 16)) + 
  theme(legend.position=c(0.1,0.9), legend.justification=c(0.1,0.9)) 

```

```{r}
ggplot(mdd, aes(x=teps_ant, y=fb_gain_rpe)) +
  geom_point(size = 5) + 
  geom_smooth(method=lm, aes(fill=group), se = FALSE) + 
  scale_color_manual(values=c('lightblue3','burlywood2')) + 
  labs(x='Anticipatory Anhedonia', y='Habenula Gain condition RPE', title='') +
  theme_light() +
  theme(axis.text = element_text(size = 16)) + 
  theme(legend.position=c(0.1,0.9), legend.justification=c(0.1,0.9)) 

```

# Anterior Insula

## Consummatory 
### gain rpe
```{r}
ins_ca <- df[df$ROI == "antrInsula", ]

hc <- ins_ca[ins_ca$group == "HC", ] # N=35
mdd <- ins_ca[ins_ca$group == "MDD", ] # N=37

# HC
res <- cor.test(hc$fb_gain_rpe, hc$teps_cons, 
                    method = "pearson")
res # p = 0.4451

# MDD
res <- cor.test(mdd$fb_gain_rpe, mdd$teps_cons, 
                    method = "pearson")
res # p = 0.3166

```

### loss rpe - MDD p = 0.15

```{r}
# HC
res <- cor.test(hc$fb_loss_rpe, hc$teps_cons, 
                    method = "pearson")
res # p = 0.8723

# MDD
res <- cor.test(mdd$fb_loss_rpe, mdd$teps_cons, 
                    method = "pearson")
res # p = 0.1469

```

## Anticipatory 
### gain rpe 
```{r}
# HC
res <- cor.test(hc$fb_gain_rpe, hc$teps_ant, 
                    method = "pearson")
res # p = 0.813

# MDD
res <- cor.test(mdd$fb_gain_rpe, mdd$teps_ant, 
                    method = "pearson")
res # p = 0.3759

```

### loss rpe

```{r}
# HC
res <- cor.test(hc$fb_loss_rpe, hc$teps_ant, 
                    method = "pearson")
res # p = 0.9422

# MDD
res <- cor.test(mdd$fb_loss_rpe, mdd$teps_ant, 
                    method = "pearson")
res # p = 0.8279

```

# sgACC

## Consummatory 
### gain rpe
```{r}
sgacc_ca <- df[df$ROI == "sgACC", ]

hc <- sgacc_ca[ins_ca$group == "HC", ] # N=35
mdd <- sgacc_ca[ins_ca$group == "MDD", ] # N=37

# HC
res <- cor.test(hc$fb_gain_rpe, hc$teps_cons, 
                    method = "pearson")
res # p = 0.2847

# MDD
res <- cor.test(mdd$fb_gain_rpe, mdd$teps_cons, 
                    method = "pearson")
res # p = 0.2709

```

### loss rpe - MDD p = 0.15

```{r}
# HC
res <- cor.test(hc$fb_loss_rpe, hc$teps_cons, 
                    method = "pearson")
res # p = 0.4089

# MDD
res <- cor.test(mdd$fb_loss_rpe, mdd$teps_cons, 
                    method = "pearson")
res # p = 0.7853

```

## Anticipatory 
### gain rpe 
```{r}
# HC
res <- cor.test(hc$fb_gain_rpe, hc$teps_ant, 
                    method = "pearson")
res # p = 0.5677

# MDD
res <- cor.test(mdd$fb_gain_rpe, mdd$teps_ant, 
                    method = "pearson")
res # p = 0.527

```

### loss rpe

```{r}
# HC
res <- cor.test(hc$fb_loss_rpe, hc$teps_ant, 
                    method = "pearson")
res # p = 0.4495

# MDD
res <- cor.test(mdd$fb_loss_rpe, mdd$teps_ant, 
                    method = "pearson")
res # p = 0.7749

```


# NAC Consummatory
HC: R = 0.13 p = 0.53
MDD: R= 0.118 p = 0.565
```{r nac stats}

nac <- df[df$roi == "rl_accumbens", ]

hc <- nac[nac$group == "HC", ]
mdd <- nac[nac$group == "MDD", ]

# the data are normally distributed -- yes, the p values are greater than 0.05 

shapiro.test(hc$feedback) # => p = 0.096
shapiro.test(hc$TEPS.Consummatory) # => p = 0.06
shapiro.test(mdd$feedback) # p =0.34
shapiro.test(mdd$TEPS.Consummatory) # p = 0.85

# HC
res <- cor.test(hc$feedback, hc$TEPS.Consummatory, 
                    method = "pearson")
res
res$p.value # 0.53

# MDD
res <- cor.test(mdd$feedback, mdd$TEPS.Consummatory, 
                    method = "pearson")
res #0.118
res$p.value # 0.565

### plot

ggplot(nac, aes(x=TEPS.Consummatory, y=feedback, color=group)) +
  geom_point(size = 5) + 
  geom_smooth(method=lm, aes(fill=group), se = FALSE) + 
  scale_color_manual(values=c('lightblue3','burlywood2')) + 
  labs(x='Consummatory Anhedonia', y='NAC Response to Feedback', title='') +
  theme_light() +
  theme(axis.text = element_text(size = 16)) + 
  theme(legend.position=c(0.1,0.9), legend.justification=c(0.1,0.9)) 

```

## NAC Anticipatory 
hc: R=0.037 p=0.858
mdd: R=0.070 p= 0.733

```{r}

# the data are normally distributed -- yes, the p values are greater than 0.05 

shapiro.test(hc$feedback) # => p = 0.096
shapiro.test(hc$TEPS.Anticipatory) # => p = 0.38
shapiro.test(mdd$feedback) # p =0.34
shapiro.test(mdd$TEPS.Anticipatory) # p = 0.997

# HC
res <- cor.test(hc$feedback, hc$TEPS.Anticipatory, 
                    method = "pearson")
res
res$p.value

# MDD
res <- cor.test(mdd$feedback, mdd$TEPS.Anticipatory, 
                    method = "pearson")
res 
res$p.value 

### plot

ggplot(nac, aes(x=TEPS.Anticipatory, y=feedback, color=group)) +
  geom_point(size = 5) + 
  geom_smooth(method=lm, aes(fill=group), se = FALSE) + 
  scale_color_manual(values=c('lightblue3','burlywood2')) + 
  labs(x='Anticipatory Anhedonia', y='NAC Response to Feedback', title='') +
  theme_light() +
  theme(axis.text = element_text(size = 16)) + 
  theme(legend.position=c(0.1,0.9), legend.justification=c(0.1,0.9)) 

```

# vmPFC Consummatory 
HC: R = -0.051, p=0.805
MDD: R=0.054, p=0.793
```{r stats}
vmpfc <- df[df$roi == "vmPFC", ]

hc <- vmpfc[vmpfc$group == "HC", ]
mdd <- vmpfc[vmpfc$group == "MDD", ]

# the data are normally distributed -- yes, the p values are greater than 0.05 

shapiro.test(hc$feedback) # => p = 0.064
shapiro.test(hc$TEPS.Consummatory) # => p = 0.06
shapiro.test(mdd$feedback) # p =0.17
shapiro.test(mdd$TEPS.Consummatory) # p = 0.85

# HC
res <- cor.test(hc$feedback, hc$TEPS.Consummatory, 
                    method = "pearson")
res
res$p.value # 0.805

# MDD
res <- cor.test(mdd$feedback, mdd$TEPS.Consummatory, 
                    method = "pearson")
res #0.054
res$p.value # 0.793

### plot

ggplot(vmpfc, aes(x=TEPS.Consummatory, y=feedback, color=group)) +
  geom_point(size = 5) + 
  geom_smooth(method=lm, aes(fill=group), se = FALSE) + 
  scale_color_manual(values=c('lightblue3','burlywood2')) + 
  labs(x='Consummatory Anhedonia', y='vmPFC Response to Feedback', title='') +
  theme_light() +
  theme(axis.text = element_text(size = 16)) + 
  theme(legend.position=c(0.1,0.9), legend.justification=c(0.1,0.9)) 

```

## vmPFC Anticipatory 

HC: R=0.00988, p = 0.962
MDD: R= -0.15 p =0.46
```{r}

# the data are normally distributed -- yes, the p values are greater than 0.05 

shapiro.test(hc$feedback) # => p = 0.064
shapiro.test(hc$TEPS.Anticipatory) # => p = 0.38
shapiro.test(mdd$feedback) # p =0.17
shapiro.test(mdd$TEPS.Anticipatory) # p = 0.997

# HC
res <- cor.test(hc$feedback, hc$TEPS.Anticipatory, 
                    method = "pearson")
res
res$p.value # 0.805

# MDD
res <- cor.test(mdd$feedback, mdd$TEPS.Anticipatory, 
                    method = "pearson")
res # -0.151
res$p.value # 0.461

### plot

ggplot(vmpfc, aes(x=TEPS.Anticipatory, y=feedback, color=group)) +
  geom_point(size = 5) + 
  geom_smooth(method=lm, aes(fill=group), se = FALSE) + 
  scale_color_manual(values=c('lightblue3','burlywood2')) + 
  labs(x='Anticipatory Anhedonia', y='vmPFC Response to Feedback', title='') +
  theme_light() +
  theme(axis.text = element_text(size = 16)) + 
  theme(legend.position=c(0.1,0.9), legend.justification=c(0.1,0.9)) 

```

#BLA 

HC: r=0.195, p=0.339
MDD: r= -0.055, p = 0.789

```{r stats}
bla_masq <- df[df$roi == "lb_amyg", ]

hc <- bla_masq[bla_masq$group == "HC", ]
mdd <- bla_masq[bla_masq$group == "MDD", ]

# the data are normally distributed -- no, p values are less than 0.05 (do spearman)

shapiro.test(hc$feedback) # => p = 1.302 * 10-5
shapiro.test(hc$masq_gd_score) # => p = 0.0034
shapiro.test(mdd$feedback) # p =0.001
shapiro.test(mdd$masq_gd_score) # p = 0.55

# HC
res <- cor.test(hc$feedback, hc$masq_gd_score, 
                    method = "spearman")
res # r = 0.195
res$p.value # 0.339

# MDD
res <- cor.test(mdd$feedback, mdd$masq_gd_score, 
                    method = "spearman")
res
res$p.value # 0.789

### plot
ggplot(bla_masq, aes(x=masq_gd_score, y=feedback, color=group)) +
  geom_point(size = 5) + 
  geom_smooth(method=lm, aes(fill=group), se = FALSE) + 
  scale_color_manual(values=c('lightblue3','burlywood2')) + 
  labs(x='General Distress', y='BLA Response to feedback', title='') +
  theme_light() +
  theme(axis.text = element_text(size = 16)) + 
  theme(legend.position=c(0.9,0.9), legend.justification=c(0.9,0.9)) 

```

# VTA MNI Consummatory
HC: R=-0.37 p=0.066
MDD: R= -0.125, p=0.54

```{r stats}

vta_consum <- df[df$roi == "vta_mni", ]

hc <- vta_consum[vta_consum$group == "HC", ]
mdd <- vta_consum[vta_consum$group == "MDD", ]

# the data are normally distributed -- yes, the p values are greater than 0.05 

shapiro.test(hc$feedback) # => p = 0.422
shapiro.test(hc$TEPS.Consummatory) # => p = 0.06
shapiro.test(mdd$feedback) # p =0.998
shapiro.test(mdd$TEPS.Consummatory) # p = 0.85

# HC
res <- cor.test(hc$feedback, hc$TEPS.Consummatory, 
                    method = "pearson")
res
res$p.value # 0.53

# MDD
res <- cor.test(mdd$feedback, mdd$TEPS.Consummatory, 
                    method = "pearson")
res
res$p.value # 0.022

### plot

ggplot(vta_consum, aes(x=TEPS.Consummatory, y=feedback, color=group)) +
  geom_point(size = 5) + 
  geom_smooth(method=lm, aes(fill=group), se = FALSE) + 
  scale_color_manual(values=c('lightblue3','burlywood2')) + 
  labs(x='Consummatory Anhedonia', y='VTA Response to Feedback', title='') +
  theme_light() +
  theme(axis.text = element_text(size = 16)) + 
  theme(legend.position=c(0.1,0.9), legend.justification=c(0.1,0.9)) 

```


HC: R = -0.36 p=0.071
MDD: R=0.127 p=0.54

```{r Anticipatory}

shapiro.test(hc$feedback) # => p = 0.422
shapiro.test(hc$TEPS.Anticipatory) # => p = 0.379
shapiro.test(mdd$feedback) # p =0.998
shapiro.test(mdd$TEPS.Anticipatory) # p = 0.997

# HC
res <- cor.test(hc$feedback, hc$TEPS.Anticipatory, 
                    method = "pearson")
res
res$p.value # 0.071

# MDD
res <- cor.test(mdd$feedback, mdd$TEPS.Anticipatory, 
                    method = "pearson")
res
res$p.value # 0.022

### plot

ggplot(vta_consum, aes(x=TEPS.Anticipatory, y=feedback, color=group)) +
  geom_point(size = 5) + 
  geom_smooth(method=lm, aes(fill=group), se = FALSE) + 
  scale_color_manual(values=c('lightblue3','burlywood2')) + 
  labs(x='Anticipatory Anhedonia', y='VTA Response to Feedback', title='') +
  theme_light() +
  theme(axis.text = element_text(size = 16)) + 
  theme(legend.position=c(0.1,0.9), legend.justification=c(0.1,0.9)) 


```

# GROUP DIFFERENCES W/T-TESTS

unpaired samples t-test is used:
1. when 2 groups are normally distributed (p > 0.05 implying distributions aren't sig. different from the normal distribution)
2. when the variances of two groups are equal (check using F test, p > 0.05 )

Ref: http://www.sthda.com/english/wiki/unpaired-two-samples-t-test-in-r

## ROI summary stats

```{r}
# select for specific regions of interest
roi_filtered <- roi %>%
  filter(
    ROI %in% c(
      "vta_mni",
      "vta_aveHCPT",
      "amyg",
      "LB_amyg",
      "vmPFC",
      "rl_accumbens",
      "antrInsula"
    )
  )

# melt just to plot / generate a summary of means & stds
roi_contrasts_m <-
  reshape2::melt(roi_filtered,
                 id = c('subj_ID', 'group', 'ROI'),
                 variable.name = "contrast")

roi_contrasts_sum <- roi_contrasts_m %>% 
  group_by(group, contrast, ROI) %>%
  summarise(mean_contrast = mean(value), sd_contrast= sd(value)) %>% arrange(contrast, ROI)

# write.csv(roi_contrasts_sum, file = "roi_summary_stats.csv", row.names=F)

```

### plot & assess for outliers

## box plot
```{r}

ggplot(roi_contrasts_m, aes(x = contrast, y = value, fill = group)) +
  geom_boxplot(outlier.shape = NA) +
    geom_dotplot(binaxis='y', stackdir='center',
                 position=position_dodge(0.9), dotsize = 0.3) +
  
  scale_fill_manual(values=c('lightblue3','peachpuff')) +

  facet_wrap(~ROI, scales = "free_y") +
  labs(title = "ROI signal intensity by contrast and group",
       x = "Contrast",
       y = "Signal") +
  theme_minimal() + 
    theme(
    # axis.text = element_text(size = 15),
    axis.title = element_text(size = 15),
    strip.text = element_text(size = 12),
    legend.title = element_blank(),
    legend.text = element_text(size=15),
  )

```

## mean + sd
```{r}
ggplot(roi_contrasts_sum, aes(x = contrast, y = mean_contrast, fill = group)) +
  geom_bar(stat = "identity", position = "dodge") +
  
  geom_errorbar(
    aes(ymin = mean_contrast - sd_contrast, ymax = mean_contrast + sd_contrast),
    width = .1,
    position = position_dodge(.9)
  ) +
  
  scale_fill_manual(values=c('lightblue3','peachpuff')) +

  facet_wrap(~ROI, scales = "free_y") +
  labs(title = "Mean ROI signal intensity by contrast and group",
       x = "Contrast",
       y = "Signal") +
  theme_minimal() + 
    theme(
    axis.text.x = element_text(angle=45, hjust=1),
    axis.title = element_text(size = 15),
    strip.text = element_text(size = 12),
    legend.title = element_blank(),
    legend.text = element_text(size=15),
  )

```

## outlier detection

```{r}

roi_out_3 <- roi_contrasts_m %>%
  group_by(group, contrast, ROI) %>%
  mutate(
    mean_value = mean(value),
    sd_value = sd(value),
    is_outlier = value > (mean_value + 3 * sd_value) |
      value < (mean_value - 3 * sd_value)
  )

# Identify outliers 
outliers_3 <- roi_out_3 %>% filter(is_outlier == T)

# View the outliers
print(outliers_3)
```

### remove outliers & generate summary

```{r}
roi_o <-
  anti_join(roi_out_3,
            outliers_3,
            by = c("subj_ID", "group", "contrast", "ROI", "value")) %>%
  select("subj_ID", "group", "ROI", "contrast", "value")

# quick check, subj 549 cg25_2, stimgain should be omitted in roi_o
roi_o %>% filter(subj_ID == 549 & ROI == "cg25_2")

```

```{r}

## summary statistics for dataset without outliers
roi_o_sum <- roi_o %>% 
  group_by(group, contrast, ROI) %>%
  summarise(mean_contrast = mean(value), sd_contrast= sd(value)) %>% arrange(contrast, ROI)

```

### replot

```{r}
ggplot(roi_o_sum, aes(x = contrast, y = mean_contrast, fill = group)) +
  geom_bar(stat = "identity", position = "dodge") +
  
  geom_errorbar(
    aes(ymin = mean_contrast - sd_contrast, ymax = mean_contrast + sd_contrast),
    width = .1,
    position = position_dodge(.9)
  ) +
  
  scale_fill_manual(values=c('lightblue3','peachpuff')) +

  facet_wrap(~ROI, scales = "free_y") +
  labs(title = "Mean ROI signal intensity by contrast and group",
       x = "Contrast",
       y = "Signal") +
  theme_minimal() + 
    theme(
    axis.text.x = element_text(angle=45, hjust=1),
    axis.title = element_text(size = 15),
    strip.text = element_text(size = 12),
    legend.title = element_blank(),
    legend.text = element_text(size=15),
  )

```

```{r unmelt}

roi_o_w <- reshape2::dcast(roi_o, subj_ID + group + ROI ~ contrast)
roi_o_w <- roi_o_w[order(roi_o_w$ROI, roi_o_w$subj_ID),]

col_order <-
  c(
    "group",
    "subj_ID",
    "stimGain",
    "stimLoss",
    "stimNeutral",
    "keyGo",
    "fb_loss_rpe",
    "fb_gain_rpe",
    "fb_all_rpe",
    "ROI"
  )

roi_o_w <- roi_o_w[, col_order]
```

# T-TESTS exploring group differences
- use roi_o_w (roi dataframe with outliers removed)

## Function
```{r function}

roi_analysis <- function(df, roi, contrast, group_col) {
  
  # subset the dataframe for the ROI and contrast of interest
  
  df_sub = df[df$ROI == roi, c(group_col, contrast)]
  
  roi_summary <- df_sub %>%
  group_by(!!rlang::sym(group_col)) %>%
    get_summary_stats(!!rlang::sym(contrast), type = "mean_sd")
  
  HC_normalcy_test <-
    with(df_sub, shapiro.test(df_sub[[contrast]][group == "HC"]))
  MDD_normalcy_test <-
    with(df_sub, shapiro.test(df_sub[[contrast]][group == "MDD"]))
  
  res.ftest <-
    var.test(df_sub[[contrast]] ~ df_sub[[group_col]])

  res_ttest <-
    t.test(df_sub[[contrast]] ~ df_sub[[group_col]])
  
  if (HC_normalcy_test$p.value < 0.05 |
      MDD_normalcy_test$p.value < 0.05 |
      res.ftest$p.value < 0.05) {
    print("Conducting non-parametric test...")
    
    wilcox_result <-
      wilcox.test(df_sub[[contrast]] ~ df_sub[[group_col]])
    t_test_result <- NULL
  } else {
    print("Conducting t-test...")
    wilcox_result <- NULL
    t_test_result <- t.test(df_sub[[contrast]] ~ df_sub[[group_col]], var.equal = T)
  }

  results <- list(
    summary_stats = roi_summary,
    HC_normalcy_test = HC_normalcy_test,
    MDD_normalcy_test = MDD_normalcy_test,
    var_test_result = res.ftest,
    wilcox_result = wilcox_result,
    t_test_result = t_test_result
  )

  return(results)
}

```


## VTA MNI
t=0.33 p=0.743

```{r}
vta_mni <- roi_o_w[roi_o_w$ROI == "vta_mni", ]

vta_mni <- vta_mni %>% select('group', 'fb_gain_rpe')

vta_fb_gain_summ <- vta_mni %>%
  group_by(group) %>%
  get_summary_stats(fb_gain_rpe, type = "mean_sd")

# test for normalcy: Data are normally distributed

with(vta_mni, shapiro.test(fb_gain_rpe[group == "HC"])) # p = 0.48
with(vta_mni, shapiro.test(fb_gain_rpe[group == "MDD"])) # p = 0.77

# variances: F=1.21 p = 0.574 ; no significant difference between the variances 
res.ftest <- var.test(fb_gain_rpe ~ group, data = vta_mni)
res.ftest

res <- t.test(fb_gain_rpe ~ group, data = vta_mni, var.equal = T) # t = 1.2134 p= 0.23
res 
```

```{r loss rpe}

vta_mni <- roi_o_w[roi_o_w$ROI == "vta_mni", ]

vta_mni <- vta_mni %>% select('group', 'fb_loss_rpe')

vta_fb_loss_summ <- vta_mni %>%
  group_by(group) %>%
  get_summary_stats(fb_loss_rpe, type = "mean_sd")

# test for normalcy: Data are normally distributed

with(vta_mni, shapiro.test(fb_loss_rpe[group == "HC"])) # p = 0.75
with(vta_mni, shapiro.test(fb_loss_rpe[group == "MDD"])) # p = 0.67

# variances: F=0.80 p = 0.521 so no significant difference between the variances 
res.ftest <- var.test(fb_loss_rpe ~ group, data = vta_mni)
res.ftest

res <- t.test(fb_loss_rpe ~ group, data = vta_mni, var.equal = T) # t = -0.68924 p= 0.493
res 

```

```{r all RPE}
vta_fb_all_summ <- vta_mni %>%
  group_by(group) %>%
  get_summary_stats(fb_all_rpe, type = "mean_sd")

with(vta_mni, shapiro.test(fb_all_rpe[group == "HC"])) # p = 0.03308
with(vta_mni, shapiro.test(fb_all_rpe[group == "MDD"])) # p = 0.2738

# variances: F=1.2991 p = 0.4341 so no significant difference between the variances 
res.ftest <- var.test(fb_all_rpe ~ group, data = vta_mni)
res.ftest

res <- t.test(fb_all_rpe ~ group, data = vta_mni, var.equal = T) # t = 0.69995 p= 0.4862
res 

wilcox.test(vta_mni$fb_all_rpe ~ vta_mni$group)
```


### plots

#### gain rpe
```{r}
ggboxplot(vta_mni, x = "group", y = "fb_all_rpe", fill = "group", add = "jitter") +
  
 scale_fill_manual(values=c('lightblue3','peachpuff')) +

  labs(x = "",
       y = "vta rpe",
       title = "") +
  
  theme_classic() +
  theme(
    axis.text = element_text(size = 20),
    axis.title = element_text(size = 24),
    axis.ticks = element_line(size = 1.5),
    legend.title = element_blank(),
    legend.position = "",
  )

```

```{r mean}
p <- ggplot(vta_fb_gain_summ, aes(x = group, y = mean, fill = group)) +
  labs(x = "",
       y = "vta gain rpe",
       title = "") +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd),
                width = .2,
                position = position_dodge(.9)) +
  scale_fill_manual(values=c('lightblue3','peachpuff')) +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 20),
    axis.title = element_text(size = 24),
    legend.title = element_blank(),
    legend.text = element_text(size=20),
    legend.position = c(0.90, 0.95))
```

#### loss rpe

```{r}
ggboxplot(vta_mni, x = "group", y = "fb_loss_rpe", fill = "group", add = "jitter") +
  
 scale_fill_manual(values=c('lightblue3','peachpuff')) +

  labs(x = "",
       y = "vta loss rpe",
       title = "") +
  
  theme_classic() +
  theme(
    axis.text = element_text(size = 20),
    axis.title = element_text(size = 24),
    axis.ticks = element_line(size = 1.5),
    legend.title = element_blank(),
    legend.position = "",
  )

p <- ggplot(vta_fb_loss_summ, aes(x = group, y = mean, fill = group)) +
  labs(x = "",
       y = "vta loss rpe",
       title = "") +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd),
                width = .2,
                position = position_dodge(.9)) +
  scale_fill_manual(values=c('lightblue3','peachpuff')) +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 20),
    axis.title = element_text(size = 24),
    legend.title = element_blank(),
    legend.text = element_text(size=20),
    legend.position = c(0.90, 0.95),
  )

```

```{r}

roi_analysis(df = roi_o_w, roi = "vta_mni", contrast = "fb_all_rpe", group = "group")
roi_analysis(df = roi_o_w, roi = "vta_mni", contrast = "fb_gain_rpe", group = "group")

roi_analysis(df = roi_o_w, roi = "vta_mni", contrast = "stimGain", group = "group")
roi_analysis(df = roi_o_w, roi = "vta_mni", contrast = "stimLoss", group = "group")


```


## vta_aveHCPT

```{r}
roi_analysis(df = roi_o_w, roi = "vta_aveHCPT", contrast = "fb_gain_rpe", group = "group")
roi_analysis(df = roi_o_w, roi = "vta_aveHCPT", contrast = "fb_loss_rpe", group = "group")
roi_analysis(df = roi_o_w, roi = "vta_aveHCPT", contrast = "fb_all_rpe", group = "group")

roi_analysis(df = roi_o_w, roi = "vta_aveHCPT", contrast = "stimGain", group = "group")
roi_analysis(df = roi_o_w, roi = "vta_aveHCPT", contrast = "stimLoss", group = "group")

```

## RL Accumbens

```{r}

roi_analysis(df = roi_o_w, roi = "rl_accumbens", contrast = "fb_gain_rpe", group = "group")
roi_analysis(df = roi_o_w, roi = "rl_accumbens", contrast = "fb_loss_rpe", group = "group")
roi_analysis(df = roi_o_w, roi = "rl_accumbens", contrast = "fb_all_rpe", group = "group")

roi_analysis(df = roi_o_w, roi = "rl_accumbens", contrast = "stimGain", group = "group")
roi_analysis(df = roi_o_w, roi = "rl_accumbens", contrast = "stimLoss", group = "group")
```

```{r}
with(nac, shapiro.test(fb_all_rpe[group == "HC"])) # p = 0.004828
with(nac, shapiro.test(fb_all_rpe[group == "MDD"])) # p = 0.8932

# variances: F=1.1395 p = 0.6949 so no significant difference between the variances 
res.ftest <- var.test(fb_all_rpe ~ group, data = nac)
res.ftest

res <- t.test(fb_all_rpe ~ group, data = nac, var.equal = T) # t = 1.8728 p= 0.06521
res 

wilcox.test(nac$fb_all_rpe ~ nac$group)

```


```{r}
nac <- roi_o_w[roi_o_w$ROI == "rl_accumbens", ]

ggboxplot(nac, x = "group", y = "fb_all_rpe", fill = "group", add = "jitter") +
  
 scale_fill_manual(values=c('lightblue3','peachpuff')) +

  labs(x = "",
       y = "nac rpe",
       title = "") +
  
  theme_classic() +
  theme(
    axis.text = element_text(size = 20),
    axis.title = element_text(size = 24),
    axis.ticks = element_line(size = 1.5),
    legend.title = element_blank(),
    legend.position = "",
  )

```


## vmPFC

```{r}
roi_analysis(df = roi_o_w, roi = "vmPFC", contrast = "fb_gain_rpe", group = "group")
roi_analysis(df = roi_o_w, roi = "vmPFC", contrast = "fb_loss_rpe", group = "group")
roi_analysis(df = roi_o_w, roi = "vmPFC", contrast = "fb_all_rpe", group = "group")

roi_analysis(df = roi_o_w, roi = "vmPFC", contrast = "stimGain", group = "group")
roi_analysis(df = roi_o_w, roi = "vmPFC", contrast = "stimLoss", group = "group")

```

### plot
```{r}

group_order <- c("HC", "MDD")

# Convert the 'group' variable to a factor with the specified order
roi_o_w$group <- factor(roi_o_w$group, levels = group_order)

vmpfc <- roi_o_w[roi_o_w$ROI == "vmPFC", ]

## gain rpe
ggboxplot(vmpfc, x = "group", y = "fb_gain_rpe", fill = "group", add = "jitter") +
  
 scale_fill_manual(values=c('lightblue3','peachpuff')) +

  labs(x = "",
       y = "vmpfc gina rpe",
       title = "") +
  
  theme_classic() +
  theme(
    axis.text = element_text(size = 20),
    axis.title = element_text(size = 24),
    axis.ticks = element_line(size = 1.5),
    legend.title = element_blank(),
    legend.position = "",
  )

## loss rpe
vmpfc <- roi_o_w[roi_o_w$ROI == "vmPFC", ]

ggboxplot(vmpfc, x = "group", y = "fb_loss_rpe", fill = "group", add = "jitter") +
  
 scale_fill_manual(values=c('lightblue3','peachpuff')) +

  labs(x = "",
       y = "vmpfc loss rpe",
       title = "") +
  
  theme_classic() +
  theme(
    axis.text = element_text(size = 20),
    axis.title = element_text(size = 24),
    axis.ticks = element_line(size = 1.5),
    legend.title = element_blank(),
    legend.position = "",
  )

# full rpe 
ggboxplot(vmpfc, x = "group", y = "fb_all_rpe", fill = "group", add = "jitter") +
  
 scale_fill_manual(values=c('lightblue3','peachpuff')) +

  labs(x = "",
       y = "vmpfc all rpe",
       title = "") +
  
  theme_classic() +
  theme(
    axis.text = element_text(size = 20),
    axis.title = element_text(size = 24),
    axis.ticks = element_line(size = 1.5),
    legend.title = element_blank(),
    legend.position = "",
  )

```


## sgACC

```{r}

sgACC <- roi_o_w[roi_o_w$ROI == "sgACC", ]


roi_analysis(df = roi_o_w, roi = "sgACC", contrast = "fb_gain_rpe", group = "group") # W = 905, p-value = 0.007951
roi_analysis(df = roi_o_w, roi = "sgACC", contrast = "fb_loss_rpe", group = "group")

roi_analysis(df = roi_o_w, roi = "sgACC", contrast = "stimGain", group = "group")
roi_analysis(df = roi_o_w, roi = "sgACC", contrast = "stimLoss", group = "group")

```

```{r}

group_order <- c("HC", "MDD")

# Convert the 'group' variable to a factor with the specified order
roi_o_w$group <- factor(roi_o_w$group, levels = group_order)

## gain rpe
ggboxplot(sgACC, x = "group", y = "fb_gain_rpe", fill = "group", add = "jitter") +
  
 scale_fill_manual(values=c('lightblue3','peachpuff')) +

  labs(x = "",
       y = "sgACC gain rpe",
       title = "") +
  
  theme_classic() +
  theme(
    axis.text = element_text(size = 20),
    axis.title = element_text(size = 24),
    axis.ticks = element_line(size = 1.5),
    legend.title = element_blank(),
    legend.position = "",
  )

## loss rpe

ggboxplot(sgACC, x = "group", y = "fb_loss_rpe", fill = "group", add = "jitter") +
  
 scale_fill_manual(values=c('lightblue3','peachpuff')) +

  labs(x = "",
       y = "sgACC loss rpe",
       title = "") +
  
  theme_classic() +
  theme(
    axis.text = element_text(size = 20),
    axis.title = element_text(size = 24),
    axis.ticks = element_line(size = 1.5),
    legend.title = element_blank(),
    legend.position = "",
  )


```


## Amygdala 

```{r}
amy <- roi_o_w[roi_o_w$ROI == "amyg", ]

roi_analysis(df = roi_o_w, roi = "amyg", contrast = "fb_all_rpe", group = "group") 
roi_analysis(df = roi_o_w, roi = "amyg", contrast = "fb_gain_rpe", group = "group") 
roi_analysis(df = roi_o_w, roi = "amyg", contrast = "fb_loss_rpe", group = "group")

roi_analysis(df = roi_o_w, roi = "amyg", contrast = "stimGain", group = "group")
roi_analysis(df = roi_o_w, roi = "amyg", contrast = "stimLoss", group = "group")

```

```{r}
group_order <- c("HC", "MDD")

# Convert the 'group' variable to a factor with the specified order
roi_o_w$group <- factor(roi_o_w$group, levels = group_order)

## gain rpe
ggboxplot(amy, x = "group", y = "fb_gain_rpe", fill = "group", add = "jitter") +
  
 scale_fill_manual(values=c('lightblue3','peachpuff')) +

  labs(x = "",
       y = "amygdala gain rpe",
       title = "") +
  
  theme_classic() +
  theme(
    axis.text = element_text(size = 20),
    axis.title = element_text(size = 24),
    axis.ticks = element_line(size = 1.5),
    legend.title = element_blank(),
    legend.position = "",
  )

## loss rpe

ggboxplot(amy, x = "group", y = "fb_loss_rpe", fill = "group", add = "jitter") +
  
 scale_fill_manual(values=c('lightblue3','peachpuff')) +

  labs(x = "",
       y = "amygdala loss rpe",
       title = "") +
  
  theme_classic() +
  theme(
    axis.text = element_text(size = 20),
    axis.title = element_text(size = 24),
    axis.ticks = element_line(size = 1.5),
    legend.title = element_blank(),
    legend.position = "",
  )

# full rpe

ggboxplot(amy, x = "group", y = "fb_all_rpe", fill = "group", add = "jitter") +
  
 scale_fill_manual(values=c('lightblue3','peachpuff')) +

  labs(x = "",
       y = "amygdala rpe",
       title = "") +
  
  theme_classic() +
  theme(
    axis.text = element_text(size = 20),
    axis.title = element_text(size = 24),
    axis.ticks = element_line(size = 1.5),
    legend.title = element_blank(),
    legend.position = "",
  )


```

## LBamyg

```{r}
lb_amyg <- roi_o_w[roi_o_w$ROI == "LB_amyg", ]

roi_analysis(df = roi_o_w, roi = "LB_amyg", contrast = "fb_all_rpe", group = "group")
roi_analysis(df = roi_o_w, roi = "LB_amyg", contrast = "fb_gain_rpe", group = "group") 
roi_analysis(df = roi_o_w, roi = "LB_amyg", contrast = "fb_loss_rpe", group = "group")

roi_analysis(df = roi_o_w, roi = "LB_amyg", contrast = "stimGain", group = "group")
roi_analysis(df = roi_o_w, roi = "LB_amyg", contrast = "stimLoss", group = "group")

```

## Ant Insula
```{r}

ant_insula <- roi_o_w[roi_o_w$ROI == "antrInsula", ]

roi_analysis(df = roi_o_w, roi = "antrInsula", contrast = "fb_all_rpe", group = "group")
roi_analysis(df = roi_o_w, roi = "antrInsula", contrast = "fb_gain_rpe", group = "group") 
roi_analysis(df = roi_o_w, roi = "antrInsula", contrast = "fb_loss_rpe", group = "group")

roi_analysis(df = roi_o_w, roi = "antrInsula", contrast = "stimGain", group = "group")
roi_analysis(df = roi_o_w, roi = "antrInsula", contrast = "stimLoss", group = "group")


```

## Habenula

```{r}

habenula <- roi_o_w[roi_o_w$ROI == "hb_mni", ]


roi_analysis(df = roi_o_w, roi = "hb_mni", contrast = "fb_gain_rpe", group = "group") 
roi_analysis(df = roi_o_w, roi = "hb_mni", contrast = "fb_loss_rpe", group = "group")

roi_analysis(df = roi_o_w, roi = "hb_mni", contrast = "stimGain", group = "group")
roi_analysis(df = roi_o_w, roi = "hb_mni", contrast = "stimLoss", group = "group")

wilcox.test(stimGain ~ group, data = habenula)
habenula %>% wilcox_effsize(stimGain ~ group)

```

## cg25

```{r}
cg25 <- roi_o_w[roi_o_w$ROI == "cg25_2", ]


roi_analysis(df = roi_o_w, roi = "cg25_2", contrast = "fb_gain_rpe", group = "group") 
roi_analysis(df = roi_o_w, roi = "cg25_2", contrast = "fb_loss_rpe", group = "group")

roi_analysis(df = roi_o_w, roi = "cg25_2", contrast = "stimGain", group = "group")
roi_analysis(df = roi_o_w, roi = "cg25_2", contrast = "stimLoss", group = "group")

```

## cg_3
```{r}

cg3 <- roi_o_w[roi_o_w$ROI == "cg_3", ]


roi_analysis(df = roi_o_w, roi = "cg_3", contrast = "fb_gain_rpe", group = "group") 
roi_analysis(df = roi_o_w, roi = "cg_3", contrast = "fb_loss_rpe", group = "group")

roi_analysis(df = roi_o_w, roi = "cg_3", contrast = "stimGain", group = "group")
roi_analysis(df = roi_o_w, roi = "cg_3", contrast = "stimLoss", group = "group")


```

### plot 
```{r}

group_order <- c("HC", "MDD")

# Convert the 'group' variable to a factor with the specified order
roi_o_w$group <- factor(roi_o_w$group, levels = group_order)

## gain rpe
ggboxplot(cg3, x = "group", y = "fb_gain_rpe", fill = "group", add = "jitter") +
  
 scale_fill_manual(values=c('lightblue3','peachpuff')) +

  labs(x = "",
       y = "cg3 gain rpe",
       title = "") +
  
  theme_classic() +
  theme(
    axis.text = element_text(size = 20),
    axis.title = element_text(size = 24),
    axis.ticks = element_line(size = 1.5),
    legend.title = element_blank(),
    legend.position = "",
  )


```

# Group analysis without outlier removal - use roi_filtered

```{r nac fb gain}

nac <- roi_filtered[roi_filtered$ROI == "rl_accumbens", ]

nac <- nac %>% select('group', 'fb_gain_rpe')

nac_fb_gain_summ <- nac %>%
  group_by(group) %>%
  get_summary_stats(fb_gain_rpe, type = "mean_sd")

# test for normalcy: Data are normally distributed

ggqqplot(nac$fb_gain_rpe)

with(nac, shapiro.test(fb_gain_rpe[group == "HC"])) # p = 0.93
with(nac, shapiro.test(fb_gain_rpe[group == "MDD"])) # p = 0.51

# variances: F=1.21 p = 0.574 ; no significant difference between the variances 
res.ftest <- var.test(fb_gain_rpe ~ group, data = nac)
res.ftest

res <- t.test(fb_gain_rpe ~ group, data = nac, var.equal = T) # t = 1.4116, df = 72, p = 0.1624
res 
```

```{r nac fb loss}
nac <- roi_filtered[roi_filtered$ROI == "rl_accumbens", ]

nac <- nac %>% select('group', 'fb_loss_rpe')

nac_fb_gain_summ <- nac %>%
  group_by(group) %>%
  get_summary_stats(fb_loss_rpe, type = "mean_sd")

# test for normalcy: Data are normally distributed

ggqqplot(nac$fb_loss_rpe)

with(nac, shapiro.test(fb_loss_rpe[group == "HC"])) # p = 0.025
with(nac, shapiro.test(fb_loss_rpe[group == "MDD"])) # p = 0.31

# variances: F=0.893 p = 0.739 ; no significant difference between the variances 
res.ftest <- var.test(fb_loss_rpe ~ group, data = nac)
res.ftest

res <- t.test(fb_loss_rpe ~ group, data = nac, var.equal = T) # t = 1.4066, df = 72, p = 0.1639
res 

wilcox.test(fb_loss_rpe ~ group, data = nac)
```

```{r nac full rpe}

nac <- roi_filtered[roi_filtered$ROI == "rl_accumbens", ]

nac <- nac %>% select('group', 'fb_all_rpe')

nac_fb_all_summ <- nac %>%
  group_by(group) %>%
  get_summary_stats(fb_all_rpe, type = "mean_sd")

# test for normalcy: Data are normally distributed

ggqqplot(nac$fb_all_rpe)

with(nac, shapiro.test(fb_all_rpe[group == "HC"])) # p = 0.004828
with(nac, shapiro.test(fb_all_rpe[group == "MDD"])) # p = 0.8932

# variances: F=1.14 p = 0.695 ; no significant difference between the variances 
res.ftest <- var.test(fb_all_rpe ~ group, data = nac)
res.ftest

res <- t.test(fb_all_rpe ~ group, data = nac, var.equal = T) # t = 1.8762, df = 72, p = 0.06469
res 

wilcox.test(fb_all_rpe ~ group, data = nac)
```

```{r vta}
roi_analysis(df = roi_filtered, roi = "vta_mni", contrast = "fb_loss_rpe", group = "group")

roi_analysis(df = roi_filtered, roi = "vta_aveHCPT", contrast = "fb_loss_rpe", group = "group")

```

```{r vmpfc}

vmpfc <- roi_filtered[roi_filtered$ROI == "vmPFC", ]

ggqqplot(vmpfc$fb_loss_rpe)
hist(vmpfc$fb_loss_rpe)

roi_analysis(df = roi_filtered, roi = "vmPFC", contrast = "fb_loss_rpe", group = "group")
t.test(fb_loss_rpe ~ group, data = vmpfc, var.equal = T)

roi_analysis(df = roi_filtered, roi = "vmPFC", contrast = "fb_gain_rpe", group = "group")
ggqqplot(vmpfc$fb_gain_rpe)
t.test(fb_gain_rpe ~ group, data = vmpfc, var.equal = T)
```

```{r antr_insula}

antr_insula <- roi_filtered[roi_filtered$ROI == "antrInsula", ]

ggqqplot(antr_insula$fb_gain_rpe)
hist(antr_insula$fb_gain_rpe)

roi_analysis(df = roi_filtered, roi = "antrInsula", contrast = "fb_gain_rpe", group = "group")

t.test(fb_gain_rpe ~ group, data = antr_insula, var.equal = T)

```

```{r amygdala}

amygdala <- roi_filtered[roi_filtered$ROI == "amyg", ]

ggqqplot(amygdala$fb_gain_rpe) # defs one outlier affecting distribution
hist(amygdala$fb_gain_rpe)

roi_analysis(df = roi_filtered, roi = "amyg", contrast = "fb_gain_rpe", group = "group")

t.test(fb_gain_rpe ~ group, data = amygdala, var.equal = T)

# all RPE
ggqqplot(amygdala$fb_all_rpe) # looks okay actually
hist(amygdala$fb_all_rpe) # outlier

roi_analysis(df = roi_filtered, roi = "amyg", contrast = "fb_all_rpe", group = "group")

t.test(fb_gain_rpe ~ group, data = amygdala, var.equal = T)

```

```{r bla}

bla <- roi_filtered[roi_filtered$ROI == "LB_amyg", ]

ggqqplot(bla$fb_gain_rpe) # defs one outlier affecting distribution
hist(bla$fb_gain_rpe)

roi_analysis(df = roi_filtered, roi = "LB_amyg", contrast = "fb_gain_rpe", group = "group")

t.test(fb_gain_rpe ~ group, data = bla, var.equal = T)

# all RPE
ggqqplot(bla$fb_all_rpe) # not great
hist(bla$fb_all_rpe) # outlier

roi_analysis(df = roi_filtered, roi = "LB_amyg", contrast = "fb_all_rpe", group = "group")

t.test(fb_all_rpe ~ group, data = bla, var.equal = F)


# stim loss
ggqqplot(bla$stimLoss) # not great
hist(bla$stimLoss) # skewed

roi_analysis(df = roi_filtered, roi = "LB_amyg", contrast = "stimLoss", group = "group")

t.test(stimLoss ~ group, data = bla, var.equal = F)


```

# Partial correlations 
- use df_covs2

```{r function}
pcor_test <- function(df, roi, contrast, scale, group_col, sex_col) {
  
  # subset the dataframe for the ROI and contrast of interest
  
  df_sub = df[df$ROI == roi,c(sex_col, group_col, contrast, scale)]
  
  HC_normalcy_contrast_test <-
    with(df_sub, shapiro.test(df_sub[[contrast]][group == "HC"]))
  
  MDD_normalcy_contrast_test <-
    with(df_sub, shapiro.test(df_sub[[contrast]][group == "MDD"]))
  
  HC_normalcy_scale_test <- with(df_sub, shapiro.test(df_sub[[scale]][group == "HC"]))
  MDD_normalcy_scale_test <- with(df_sub, shapiro.test(df_sub[[scale]][group == "MDD"]))

  if (HC_normalcy_contrast_test$p.value < 0.05 |
      HC_normalcy_scale_test$p.value < 0.05) {
    print("Conducting non-parametric test...")
    hc_spearman_result <-
      pcor.test(df_sub[[contrast]][df_sub[[group_col]] == "HC"],
               df_sub[[scale]][df_sub[[group_col]] == "HC"],
               df_sub[[sex_col]][df_sub[[group_col]] == "HC"],
               method = "spearman")
    hc_pearson_result <- NULL
    
  } else {
    print("Conducting pearson correlation test...")
    hc_spearman_result <- NULL
    hc_pearson_result <-
      pcor.test(df_sub[[contrast]][df_sub[[group_col]] == "HC"],
               df_sub[[scale]][df_sub[[group_col]] == "HC"],
               df_sub[[sex_col]][df_sub[[group_col]] == "HC"],
               method = "pearson")
  }
  
  if (MDD_normalcy_contrast_test$p.value < 0.05 |
    MDD_normalcy_scale_test$p.value < 0.05) {
    print("Conducting non-parametric test...")
    
    mdd_spearman_result <-
      pcor.test(df_sub[[contrast]][df_sub[[group_col]] == "MDD"],
               df_sub[[scale]][df_sub[[group_col]] == "MDD"],
               df_sub[[sex_col]][df_sub[[group_col]] == "MDD"],
               method = "spearman")
    mdd_pearson_result <- NULL
  } else {
    print("Conducting pearson correlation test...")
    mdd_spearman_result <- NULL
    mdd_pearson_result <-
      pcor.test(df_sub[[contrast]][df_sub[[group_col]] == "MDD"],
               df_sub[[scale]][df_sub[[group_col]] == "MDD"], 
               df_sub[[sex_col]][df_sub[[group_col]] == "MDD"],
               method = "pearson")
  }

  results <- list(
    HC_normalcy_contrast_test = HC_normalcy_contrast_test,
    HC_normalcy_scale_test = HC_normalcy_scale_test,
    MDD_normalcy_contrast_test = MDD_normalcy_contrast_test,
    MDD_normalcy_scale_test = MDD_normalcy_scale_test,
    hc_spearman_result = hc_spearman_result,
    hc_pearson_result = hc_pearson_result,
    mdd_spearman_result = mdd_spearman_result,
    mdd_pearson_result = mdd_pearson_result
  )

  return(results)
}

```

```{r}

df_covs <- df_covs %>% mutate(SUB_gender=recode(SUB_gender, 'F'='1', 'M'='0')) # dummy code sex
df_covs <- df_covs %>% mutate(Medication=recode(Medication, '0'='1', '1'='2')) # 1 = none, 2 = y
df_covs$SUB_gender <- as.numeric(df_covs$SUB_gender)
df_covs$Medication <- as.numeric(df_covs$Medication)
df_covs2 <- na.omit(df_covs) # remove nan

```

# VTA
## test controlling for sex
```{r vta mni - consummatory}

# fb all rpe: HC -1.796, p=0.082, MDD: -0.1136 p 0.5156
pcor_test(
  df = df_covs2,
  roi = "vta_mni",
  contrast = "fb_all_rpe",
  scale = "TEPS.Consummatory",
  group_col = "group",
  sex_col = "SUB_gender"
)

# fb gain rpe: HC -0.0881, p 0.6258, MDD: 0.0581, p 0.740
pcor_test(
  df = df_covs2,
  roi = "vta_mni",
  contrast = "fb_gain_rpe",
  scale = "TEPS.Consummatory",
  group_col = "group",
  sex_col = "SUB_gender"
)

# fb loss rpe: HC -0.2876 p 0.1045, MDD: -0.181, p 0.297

pcor_test(
  df = df_covs2,
  roi = "vta_mni",
  contrast = "fb_loss_rpe",
  scale = "TEPS.Consummatory",
  group_col = "group",
  sex_col = "SUB_gender"
)
```

```{r vta_mni anticipatory}
# fb all rpe: HC -0.0303, p=0.8668, MDD: -0.0496 p 0.777
pcor_test(
  df = df_covs2,
  roi = "vta_mni",
  contrast = "fb_all_rpe",
  scale = "TEPS.Anticipatory",
  group_col = "group",
  sex_col = "SUB_gender"
)

# fb gain rpe: HC 0.0621, p 0.731, MDD: 0.1109, p 0.5256
pcor_test(
  df = df_covs2,
  roi = "vta_mni",
  contrast = "fb_gain_rpe",
  scale = "TEPS.Anticipatory",
  group_col = "group",
  sex_col = "SUB_gender"
)

# fb loss rpe: HC -0.092 p 0.610475, MDD: -0.1518914, p 0.3837343

pcor_test(
  df = df_covs2,
  roi = "vta_mni",
  contrast = "fb_loss_rpe",
  scale = "TEPS.Anticipatory",
  group_col = "group",
  sex_col = "SUB_gender"
)
```

## test controlling for age

```{r ant}
# fb all rpe: HC -0.3196058, p=0.06981793, MDD: -0.0910885 p 0.6027849
pcor_test(
  df = df_covs2,
  roi = "vta_mni",
  contrast = "fb_all_rpe",
  scale = "TEPS.Consummatory",
  group_col = "group",
  sex_col = "SUB_Age"
)

# fb gain rpe: HC -0.1036987, p 0.565774, MDD: 0.06144647, p 0.7258497
pcor_test(
  df = df_covs2,
  roi = "vta_mni",
  contrast = "fb_gain_rpe",
  scale = "TEPS.Consummatory",
  group_col = "group",
  sex_col = "SUB_Age"
)

# fb loss rpe: HC -0.2940613 p 0.0966983, MDD: -0.1586564, p 0.3626537

pcor_test(
  df = df_covs2,
  roi = "vta_mni",
  contrast = "fb_loss_rpe",
  scale = "TEPS.Consummatory",
  group_col = "group",
  sex_col = "SUB_Age"
)
```

```{r vta anticipatory}
# fb all rpe: HC 
pcor_test(
  df = df_covs2,
  roi = "vta_mni",
  contrast = "fb_all_rpe",
  scale = "TEPS.Anticipatory",
  group_col = "group",
  sex_col = "SUB_Age"
)

# fb gain rpe: HC
pcor_test(
  df = df_covs2,
  roi = "vta_mni",
  contrast = "fb_gain_rpe",
  scale = "TEPS.Anticipatory",
  group_col = "group",
  sex_col = "SUB_Age"
)

# fb loss rpe:

pcor_test(
  df = df_covs2,
  roi = "vta_mni",
  contrast = "fb_loss_rpe",
  scale = "TEPS.Anticipatory",
  group_col = "group",
  sex_col = "SUB_Age"
)

```

## test controlling for medication 

```{r}
# fb all rpe: HC -0.3196058, p=0.06981793, MDD: -0.0910885 p 0.6027849
pcor_test(
  df = df_covs2,
  roi = "vta_mni",
  contrast = "fb_all_rpe",
  scale = "TEPS.Consummatory",
  group_col = "group",
  sex_col = "Medication"
)

# fb gain rpe: HC -0.1036987, p 0.565774, MDD: 0.06144647, p 0.7258497
pcor_test(
  df = df_covs2,
  roi = "vta_mni",
  contrast = "fb_gain_rpe",
  scale = "TEPS.Consummatory",
  group_col = "group",
  sex_col = "Medication"
)

# fb loss rpe: HC -0.2940613 p 0.0966983, MDD: -0.1586564, p 0.3626537

pcor_test(
  df = df_covs2,
  roi = "vta_mni",
  contrast = "fb_loss_rpe",
  scale = "TEPS.Consummatory",
  group_col = "group",
  sex_col = "Medication"
)
```

## NAC
## test controlling for sex
```{r nac - consummatory}

# fb all rpe: HC 0.2595317, p=0.138275, MDD: 0.1287821 p 0.4609434
pcor_test(
  df = df_covs2,
  roi = "rl_accumbens",
  contrast = "fb_all_rpe",
  scale = "TEPS.Consummatory",
  group_col = "group",
  sex_col = "SUB_gender"
)

# fb gain rpe: HC 0.3163341, p 0.06835482, MDD: -0.1218136, p 0.4857471
pcor_test(
  df = df_covs2,
  roi = "rl_accumbens",
  contrast = "fb_gain_rpe",
  scale = "TEPS.Consummatory",
  group_col = "group",
  sex_col = "SUB_gender"
)

# fb loss rpe: HC 0.1319101 p 0.4570882, MDD: 0.2555815, p 0.138376

pcor_test(
  df = df_covs2,
  roi = "rl_accumbens",
  contrast = "fb_loss_rpe",
  scale = "TEPS.Consummatory",
  group_col = "group",
  sex_col = "SUB_gender"
)
```

```{r nac anticipatory}
# fb all rpe: HC 0.06484988, p=0.7155768, MDD: 0.2118275 p 0.2218601
pcor_test(
  df = df_covs2,
  roi = "rl_accumbens",
  contrast = "fb_all_rpe",
  scale = "TEPS.Anticipatory",
  group_col = "group",
  sex_col = "SUB_gender"
)

# fb gain rpe: HC 0.1712651, p 0.3328119, MDD: -0.06345877, p 0.7172339
pcor_test(
  df = df_covs2,
  roi = "rl_accumbens",
  contrast = "fb_gain_rpe",
  scale = "TEPS.Anticipatory",
  group_col = "group",
  sex_col = "SUB_gender"
)

# fb loss rpe: HC -0.1269058 p 0.4744836, MDD: 0.3175233, p 0.06306993

pcor_test(
  df = df_covs2,
  roi = "rl_accumbens",
  contrast = "fb_loss_rpe",
  scale = "TEPS.Anticipatory",
  group_col = "group",
  sex_col = "SUB_gender"
)
```

## test controlling for age

```{r ant}
# fb all rpe: HC 0.2338343, p=0.1831835, MDD: 0.02592091 p 0.8824962
pcor_test(
  df = df_covs2,
  roi = "rl_accumbens",
  contrast = "fb_all_rpe",
  scale = "TEPS.Consummatory",
  group_col = "group",
  sex_col = "SUB_Age"
)

# fb gain rpe: HC 0.291846, p 0.09397908, MDD: -0.2005064, p 0.2481285
pcor_test(
  df = df_covs2,
  roi = "rl_accumbens",
  contrast = "fb_gain_rpe",
  scale = "TEPS.Consummatory",
  group_col = "group",
  sex_col = "SUB_Age"
)

# fb loss rpe: HC 0.09807015 p 0.581095, MDD: 0.2006768, p 0.2477184

pcor_test(
  df = df_covs2,
  roi = "rl_accumbens",
  contrast = "fb_loss_rpe",
  scale = "TEPS.Consummatory",
  group_col = "group",
  sex_col = "SUB_Age"
)
```

```{r vta anticipatory}
# fb all rpe: HC: 0.06839628 p 0.70072 ; MDD 0.123111 p 0.4810775
pcor_test(
  df = df_covs2,
  roi = "rl_accumbens",
  contrast = "fb_all_rpe",
  scale = "TEPS.Anticipatory",
  group_col = "group",
  sex_col = "SUB_Age"
)

# fb gain rpe: HC: 0.187625 p 0.2879708 ; MDD -0.104979, p 0.5483925
pcor_test(
  df = df_covs2,
  roi = "rl_accumbens",
  contrast = "fb_gain_rpe",
  scale = "TEPS.Anticipatory",
  group_col = "group",
  sex_col = "SUB_Age"
)

# fb loss rpe: HC: -0.1385585, p 0.4345153 ; MDD: 0.2573208, p 0.1356197

pcor_test(
  df = df_covs2,
  roi = "rl_accumbens",
  contrast = "fb_loss_rpe",
  scale = "TEPS.Anticipatory",
  group_col = "group",
  sex_col = "SUB_Age"
)

```

## test controlling for medication 

```{r}
# fb all rpe: warning
pcor_test(
  df = df_covs2,
  roi = "rl_accumbens",
  contrast = "fb_all_rpe",
  scale = "TEPS.Consummatory",
  group_col = "group",
  sex_col = "Medication"
)

# fb gain rpe: warning
pcor_test(
  df = df_covs2,
  roi = "rl_accumbens",
  contrast = "fb_gain_rpe",
  scale = "TEPS.Consummatory",
  group_col = "group",
  sex_col = "Medication"
)

# fb loss rpe: warning

pcor_test(
  df = df_covs2,
  roi = "rl_accumbens",
  contrast = "fb_loss_rpe",
  scale = "TEPS.Consummatory",
  group_col = "group",
  sex_col = "Medication"
)
```

## vmPFC
## test controlling for sex
```{r vmpfc - consummatory}

# fb all rpe: HC 0.05333303, p=0.7681685, MDD: 0.1704553 p 0.3275835
pcor_test(
  df = df_covs2,
  roi = "vmPFC",
  contrast = "fb_all_rpe",
  scale = "TEPS.Consummatory",
  group_col = "group",
  sex_col = "SUB_gender"
)

# fb gain rpe: HC 0.2149226, p 0.2297012, MDD: 0.2364913, p 0.1713799
pcor_test(
  df = df_covs2,
  roi = "vmPFC",
  contrast = "fb_gain_rpe",
  scale = "TEPS.Consummatory",
  group_col = "group",
  sex_col = "SUB_gender"
)

# fb loss rpe: HC -0.0739292 p 0.6826308, MDD: -0.0008511212, p 0.9961283

pcor_test(
  df = df_covs2,
  roi = "vmPFC",
  contrast = "fb_loss_rpe",
  scale = "TEPS.Consummatory",
  group_col = "group",
  sex_col = "SUB_gender"
)
```

```{r nac anticipatory}
# fb all rpe: HC -0.1600022, p=0.3737603, MDD: 0.2105215 p 0.2247894
pcor_test(
  df = df_covs2,
  roi = "vmPFC",
  contrast = "fb_all_rpe",
  scale = "TEPS.Anticipatory",
  group_col = "group",
  sex_col = "SUB_gender"
)

# fb gain rpe: HC -0.006511016, p 0.9713134, MDD: 0.3677861, p 0.02973073 SIG!
pcor_test(
  df = df_covs2,
  roi = "vmPFC",
  contrast = "fb_gain_rpe",
  scale = "TEPS.Anticipatory",
  group_col = "group",
  sex_col = "SUB_gender"
)

# fb loss rpe: HC -0.143508 p 0.4256017, MDD: -0.07066256, p 0.6866776

pcor_test(
  df = df_covs2,
  roi = "vmPFC",
  contrast = "fb_loss_rpe",
  scale = "TEPS.Anticipatory",
  group_col = "group",
  sex_col = "SUB_gender"
)
```

## test controlling for age

```{r cons}
# fb all rpe: HC 0.05747456, p=0.7507138, MDD: 0.1230129 p 0.4814299
pcor_test(
  df = df_covs2,
  roi = "vmPFC",
  contrast = "fb_all_rpe",
  scale = "TEPS.Consummatory",
  group_col = "group",
  sex_col = "SUB_Age"
)

# fb gain rpe: HC 0.2017139, p 0.2602937, MDD: 0.2115004, p 0.2225913
pcor_test(
  df = df_covs2,
  roi = "vmPFC",
  contrast = "fb_gain_rpe",
  scale = "TEPS.Consummatory",
  group_col = "group",
  sex_col = "SUB_Age"
)

# fb loss rpe: HC -0.05086183 p 0.7786376, MDD: -0.03556113, p 0.8392867

pcor_test(
  df = df_covs2,
  roi = "vmPFC",
  contrast = "fb_loss_rpe",
  scale = "TEPS.Consummatory",
  group_col = "group",
  sex_col = "SUB_Age"
)
```

```{r vta anticipatory}
# fb all rpe: HC: -0.1629218 p 0.3649883 ; MDD 0.179149 p 0.3031372
pcor_test(
  df = df_covs2,
  roi = "vmPFC",
  contrast = "fb_all_rpe",
  scale = "TEPS.Anticipatory",
  group_col = "group",
  sex_col = "SUB_Age"
)

# fb gain rpe: HC: -0.03282482 p 0.856099 ; MDD 0.3525793, p 0.0377672 # SIG!
pcor_test(
  df = df_covs2,
  roi = "vmPFC",
  contrast = "fb_gain_rpe",
  scale = "TEPS.Anticipatory",
  group_col = "group",
  sex_col = "SUB_Age"
)

# fb loss rpe: HC: -0.1317729, p 0.4647873 ; MDD: -0.09260031, p 0.596754

pcor_test(
  df = df_covs2,
  roi = "vmPFC",
  contrast = "fb_loss_rpe",
  scale = "TEPS.Anticipatory",
  group_col = "group",
  sex_col = "SUB_Age"
)

```

## test controlling for medication 

```{r}
# fb all rpe: warning
pcor_test(
  df = df_covs2,
  roi = "vmPFC",
  contrast = "fb_all_rpe",
  scale = "TEPS.Consummatory",
  group_col = "group",
  sex_col = "Medication"
)

# fb gain rpe: warning
pcor_test(
  df = df_covs2,
  roi = "vmPFC",
  contrast = "fb_gain_rpe",
  scale = "TEPS.Consummatory",
  group_col = "group",
  sex_col = "Medication"
)

# fb loss rpe: warning

pcor_test(
  df = df_covs2,
  roi = "vmPFC",
  contrast = "fb_loss_rpe",
  scale = "TEPS.Consummatory",
  group_col = "group",
  sex_col = "Medication"
)
```


# omitting medication subjects analysis

Mean differences - none 
this is "cheeky" but for my own curiosity 
```{r vta}

roi_analysis(df = df_covs_nomed, roi = "vta_mni", contrast = "fb_all_rpe", group = "group") # p = 0.9333

roi_analysis(df = df_covs_nomed, roi = "vta_mni", contrast = "fb_gain_rpe", group = "group") # p = 0.2288

roi_analysis(df = df_covs_nomed, roi = "vta_mni", contrast = "fb_loss_rpe", group = "group") # p = 0.6633

```

correlation with teps
```{r}

ca_analysis(df = df_covs_nomed, roi = "vta_mni", contrast = "fb_all_rpe", 
            scale = "TEPS.Consummatory", group_col = "group") # this reqs spearman, MDD: p=0.7848 

ca_analysis(df = df_covs_nomed, roi = "vta_mni", contrast = "fb_gain_rpe", 
            scale = "TEPS.Consummatory", group_col = "group") # MDD: p = 0.1941, R = 0.321

ca_analysis(df = df_covs_nomed, roi = "vta_mni", contrast = "fb_loss_rpe", 
            scale = "TEPS.Consummatory", group_col = "group") # MDD: p = .563, R = -0.1460791

```



#testing function works!
- even when controlling for age, sex, medication -- no correlation with VTA signal
```{r testing function works!}

vta_sub <- df_covs2 %>% filter(ROI == "vta_mni" & group == "MDD") 

res <-
  pcor.test(vta_sub$TEPS.Consummatory,
            vta_sub$fb_all_rpe,
            vta_sub$SUB_gender,
            method = "pearson")
res$estimate

## Consummatory
# R: -0.1127149 p: 0.5322819
pcor.test(
  vta_sub$TEPS.Consummatory,
  vta_sub$fb_all_rpe,
  list(vta_sub$SUB_gender, vta_sub$SUB_Age, vta_sub$Medication)
  )

# R: 0.08036353 p: 0.6566284
pcor.test(
  vta_sub$TEPS.Consummatory,
  vta_sub$fb_gain_rpe,
  list(vta_sub$SUB_gender, vta_sub$SUB_Age, vta_sub$Medication)
  )

# R: -0.1999885 p: 0.2644787
pcor.test(
  vta_sub$TEPS.Consummatory,
  vta_sub$fb_loss_rpe,
  list(vta_sub$SUB_gender, vta_sub$SUB_Age, vta_sub$Medication)
  )

## Anticipatory 
# R: -0.0374181 p: 0.8362171
pcor.test(
  vta_sub$TEPS.Anticipatory,
  vta_sub$fb_all_rpe,
  list(vta_sub$SUB_gender, vta_sub$SUB_Age, vta_sub$Medication)
  )

# R: 0.1922603 p: 0.2837625
pcor.test(
  vta_sub$TEPS.Anticipatory,
  vta_sub$fb_gain_rpe,
  list(vta_sub$SUB_gender, vta_sub$SUB_Age, vta_sub$Medication)
  )

# R: -0.2010764 p: 0.2618349
pcor.test(
  vta_sub$TEPS.Anticipatory,
  vta_sub$fb_loss_rpe,
  list(vta_sub$SUB_gender, vta_sub$SUB_Age, vta_sub$Medication)
)

```

```{r nac}
nac_sub <- df_covs2 %>% filter(ROI == "rl_accumbens" & group == "MDD") 

## Consummatory
# R: 0.1174166 p: 0.5152038
pcor.test(
  nac_sub$TEPS.Consummatory,
  nac_sub$fb_all_rpe,
  list(nac_sub$SUB_gender, nac_sub$SUB_Age, nac_sub$Medication)
  )

# R: -0.1475583 p: 0.4125162
pcor.test(
  nac_sub$TEPS.Consummatory,
  nac_sub$fb_gain_rpe,
  list(nac_sub$SUB_gender, nac_sub$SUB_Age, nac_sub$Medication)
  )

# R: 0.2522138 p: 0.1567744 -- only trending result if we can even call it that
pcor.test(
  nac_sub$TEPS.Consummatory,
  nac_sub$fb_loss_rpe,
  list(nac_sub$SUB_gender, nac_sub$SUB_Age, nac_sub$Medication)
  )

## Anticipatory 
# R: 0.1714182 p: 0.3401652
pcor.test(
  nac_sub$TEPS.Anticipatory,
  nac_sub$fb_all_rpe,
  list(nac_sub$SUB_gender, nac_sub$SUB_Age, nac_sub$Medication)
  )

# R: -0.1097372 p: 0.5432365
pcor.test(
  nac_sub$TEPS.Anticipatory,
  nac_sub$fb_gain_rpe,
  list(nac_sub$SUB_gender, nac_sub$SUB_Age, nac_sub$Medication)
  )

# R: 0.294399 p: 0.096298
pcor.test(
  nac_sub$TEPS.Anticipatory,
  nac_sub$fb_loss_rpe,
  list(nac_sub$SUB_gender, nac_sub$SUB_Age, nac_sub$Medication)
  )

```


```{r vmpfc}
vmpfc_sub <- df_covs2 %>% filter(ROI == "vmPFC" & group == "MDD") 

## Consummatory
# R: 0.1657239 p: 0.3566856
pcor.test(
  vmpfc_sub$TEPS.Consummatory,
  vmpfc_sub$fb_all_rpe,
  list(vmpfc_sub$SUB_gender, vmpfc_sub$SUB_Age, vmpfc_sub$Medication)
  )

# R: 0.2412591 p: 0.1761937
pcor.test(
  vmpfc_sub$TEPS.Consummatory,
  vmpfc_sub$fb_gain_rpe,
  list(vmpfc_sub$SUB_gender, vmpfc_sub$SUB_Age, vmpfc_sub$Medication)
  )

# R: -0.005566784 p: 0.9754722
pcor.test(
  vmpfc_sub$TEPS.Consummatory,
  vmpfc_sub$fb_loss_rpe,
  list(vmpfc_sub$SUB_gender, vmpfc_sub$SUB_Age, vmpfc_sub$Medication)
  )

## Anticipatory 
# R: 0.2009763 p: 0.2620774
pcor.test(
  vmpfc_sub$TEPS.Anticipatory,
  vmpfc_sub$fb_all_rpe,
  list(vmpfc_sub$SUB_gender, vmpfc_sub$SUB_Age, vmpfc_sub$Medication)
  )

# R: 0.3649113 p: 0.036794
pcor.test(
  vmpfc_sub$TEPS.Anticipatory,
  vmpfc_sub$fb_gain_rpe,
  list(vmpfc_sub$SUB_gender, vmpfc_sub$SUB_Age, vmpfc_sub$Medication)
  )

# R: -0.07141358 p: 0.6928971
pcor.test(
  vmpfc_sub$TEPS.Anticipatory,
  vmpfc_sub$fb_loss_rpe,
  list(vmpfc_sub$SUB_gender, vmpfc_sub$SUB_Age, vmpfc_sub$Medication)
  )

```

# Regressions controlling for covariates
- like t-test but with covariates included
roi_o_w = 518 observations and used for t-test 

```{r}
roi_covs <- merge(roi_o_w, covs)
length(unique(roi_covs$subj_ID)) # N = 74
```

```{r vta }

vta_test <- roi_covs %>% filter(ROI == "vta_mni") 

# FB GAIN
test <- lm(vta_test$fb_gain_rpe ~ vta_test$group) # p = 0.228, t = -1.216, same as t-test
summary(test)

# what does adding sex do to these results? makes the relationship weaker
test <- lm(vta_test$fb_gain_rpe ~ vta_test$group + vta_test$SUB_gender) # p = 0.292, t = -1.062
summary(test)

# add age
test <- lm(vta_test$fb_gain_rpe ~ vta_test$group + vta_test$SUB_gender + vta_test$SUB_Age) # p = 0.317, t = -1.007
summary(test)

# add medication 
test <-
  lm(
    vta_test$fb_gain_rpe ~ vta_test$group + vta_test$SUB_gender + vta_test$SUB_Age +
      vta_test$Medication
  ) # p = 0.201, t = -1.291
summary(test)

# FB LOSS
test <- lm(vta_test$fb_loss_rpe ~ vta_test$group + vta_test$SUB_gender) # p = 0.521, t = 0.646
summary(test)

test <- lm(vta_test$fb_loss_rpe ~ vta_test$group + vta_test$SUB_gender + vta_test$SUB_Age) 
summary(test) # p = 0.482, t = 0.708

test <- lm(vta_test$fb_loss_rpe ~ vta_test$group + vta_test$SUB_gender + vta_test$SUB_Age
           + vta_test$Medication) 
summary(test) # p = 0.643, t = 0.465

# FB FULL
test <- lm(vta_test$fb_all_rpe ~ vta_test$group + vta_test$SUB_gender) 
summary(test) # p = 0.572, t = -0.567

test <- lm(vta_test$fb_all_rpe ~ vta_test$group + vta_test$SUB_gender + vta_test$SUB_Age) 
summary(test) # p = 0.650, t = -0.455

test <- lm(vta_test$fb_all_rpe ~ vta_test$group + vta_test$SUB_gender + vta_test$SUB_Age
           + vta_test$Medication) 
summary(test) # p = 0.406, t = -0.835


```

```{r nac}
nac_test <- roi_covs %>% filter(ROI == "rl_accumbens") 

# FB GAIN
test <- lm(nac_test$fb_gain_rpe ~ nac_test$group) # p = 0.16237, t = -1.412, same as t-test
summary(test)

# what does adding sex do to these results? makes the relationship weaker
test <- lm(nac_test$fb_gain_rpe ~ nac_test$group + nac_test$SUB_gender) # p = 0.2483, t = -1.164
summary(test)

# add age
test <- lm(nac_test$fb_gain_rpe ~ nac_test$group + nac_test$SUB_gender + nac_test$SUB_Age) # p = 0.2024, t = -1.287
summary(test)

# add medication 
test <-
  lm(
    nac_test$fb_gain_rpe ~ nac_test$group + nac_test$SUB_gender + nac_test$SUB_Age +
      nac_test$Medication
  ) 
summary(test) # p = 0.9657, t = 0.043 gender is sig. and medication

# FB LOSS
test <- lm(nac_test$fb_loss_rpe ~ nac_test$group + nac_test$SUB_gender) # p = 0.2035, t = -1.283
summary(test)

test <- lm(nac_test$fb_loss_rpe ~ nac_test$group + nac_test$SUB_gender + nac_test$SUB_Age) 
summary(test) # p = .1446, t = -1.475

test <- lm(nac_test$fb_loss_rpe ~ nac_test$group + nac_test$SUB_gender + nac_test$SUB_Age
           + nac_test$Medication) 
summary(test) # p = 0.5638, t = -0.580

# FB FULL
test <- lm(nac_test$fb_all_rpe ~ nac_test$group + nac_test$SUB_gender) 
summary(test) # p = 0.1034, t = -1.650

test <- lm(nac_test$fb_all_rpe ~ nac_test$group + nac_test$SUB_gender + nac_test$SUB_Age) 
summary(test) # p = 0.0642, t = -1.880 # group is trending, with sex as a significant predictor

test <- lm(nac_test$fb_all_rpe ~ nac_test$group + nac_test$SUB_gender + nac_test$SUB_Age
           + nac_test$Medication) 
summary(test) # p = 0.68077, t = -0.413, sex and meds are sig

```

# Are RPEs from different regions correlated?
- this is inspired by 2018 RPE paper

```{r}
vta_rpe <- roi_o_w %>% filter(ROI == "vta_mni" & group == "MDD") %>% dplyr::select('fb_gain_rpe')
names(vta_rpe)[names(vta_rpe) == "fb_gain_rpe"] <- "vta_gain_rpe"

nac_rpe <- roi_o_w %>% filter(ROI == "rl_accumbens" & group == "MDD") %>%
  dplyr::select('fb_gain_rpe')
names(nac_rpe)[names(nac_rpe) == "fb_gain_rpe"] <- "nac_gain_rpe"

rpe_df <- cbind(nac_rpe, vta_rpe)
```


```{r}

# as VTA loss RPE signal gets stronger, so does the NAC loss RPE in MDD; p-value = 0.03371

ggplot(rpe_df, aes(x=vta_gain_rpe, y=nac_gain_rpe)) +
  geom_point(size = 5) + 
  geom_smooth(method=lm, se = FALSE) + 
  scale_color_manual(values=c('burlywood2')) + 
  labs(x='VTA gain RPE', y='NAC gain RPE', title='') +
  theme_light() +
  theme(axis.text = element_text(size = 16)) + 
  theme(legend.position=c(0.1,0.9), legend.justification=c(0.1,0.9)) 

cor.test(rpe_df$vta_gain_rpe, rpe_df$nac_gain_rpe)

```

```{r}

```

Exclusion for < 50% correct responses
https://www.nature.com/articles/s41386-018-0032-x#Sec11
