---
title: "RPLearning_modeling"
author: "Jackie Beltrán"
date: "2023-06-26"
output: html_document
editor_options: 
  chunk_output_type: console
---

# source 
http://www.sthda.com/english/wiki/unpaired-two-samples-t-test-in-r

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(ggplot2)
require(tidyr)
require(dplyr)
require(ggpubr)
require(data.table)

require(tidyverse)
require(rstatix)
```

```{r load data}

getwd()

data_m2 <- read.csv('/Users/jackiebeltran/Documents/GitHub/RP_Learning/3_results/model_2/N76_parameters.csv')

# add group information
lrs <-
  data_m2 %>% select('subject_ID',
                     'fit_reward_alpha',
                     'fit_punishment_alpha',
                     'fit_beta') %>% mutate(group = ifelse(substr(subject_ID, 1, 1) == "4", "MDD", "HC"))



```


# Are the average reward learning rates different between HC and MDD?

HC: 0.334(0.352)
MDD: 0.307(0.370)

https://www.datanovia.com/en/lessons/wilcoxon-test-in-r/#computation-1

```{r independent t test, reward alpha}
# Summary statistics
lrs %>%
  group_by(group) %>%
  get_summary_stats(fit_reward_alpha, type = "mean_sd")

# test for normalcy: Data are not normally distributed, switch to wilcoxon test

with(lrs, shapiro.test(fit_reward_alpha[group == "HC"]))# p = 0.004693
with(lrs, shapiro.test(fit_reward_alpha[group == "MDD"])) # p = 0.000933

# variances: F=0.948 p = 0.87 so no significant difference between the variances 
res.ftest <- var.test(fit_reward_alpha ~ group, data = lrs)
res.ftest

res <- t.test(fit_reward_alpha ~ group, data = lrs) # t = 0.677 p= 0.50
res

# using an unpaired two-samples Wilcoxon test, we conclude there is no significant difference in learning rates between HC and MDD (W=807, p=0.38)

# wilcox test
stat.test <- lrs %>% 
  wilcox_test(fit_reward_alpha ~ group) %>%
  add_significance()
stat.test

# calculate effect size
lrs %>% wilcox_effsize(fit_reward_alpha ~ group) # small effect size = 0.101

```


```{r t test, beta}
# Summary statistics
lrs %>%
  group_by(group) %>%
  get_summary_stats(fit_beta, type = "mean_sd")

# test for normalcy: Data are not normally distributed, switch to wilcoxon test

with(lrs, shapiro.test(fit_beta[group == "HC"]))# p = 6.23e-07
with(lrs, shapiro.test(fit_beta[group == "MDD"])) # p = 1.18*e-6

# variances: F=0.964 p = 0.91 so no significant difference between the variances bc p>0.05
res.ftest <- var.test(fit_beta ~ group, data = lrs)
res.ftest

res <- t.test(fit_beta ~ group, data = lrs) # t = -0.068 p= 0.95
res

# using an unpaired two-samples Wilcoxon test, we conclude there is no significant difference in beta between HC and MDD (W=764, p=0.664)
# wilcox.test(fit_beta ~ group, data = lrs)


# wilcox test
stat.beta.test <- lrs %>% 
  wilcox_test(fit_beta ~ group) %>%
  add_significance()
stat.beta.test

# calculate effect size
lrs %>% wilcox_effsize(fit_beta ~ group) # small effect size = 0.051

```


# Correlation of model 1 learning rate with model 2 punishment
```{r parameter recovery correlation}
alphas <- param_recovery %>% select('alpha', 'fit_alpha')

ggscatter(alphas, x = "alpha", y = "fit_alpha", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "sim alpha", ylab = "fit alpha")

# is the data normally distributed-- no, the p values are less than 0.05 -- switch to non-parametric test

shapiro.test(alphas$alpha) # => p = 0.0061
# Shapiro-Wilk normality test for wt
shapiro.test(alphas$fit_alpha) # => p = 0.0203

# test: S = 39106, p < 2.2e-16

res <- cor.test(alphas$alpha, alphas$fit_alpha, 
                    method = "spearman")

res$p.value # 1.84 * 10^-20

plot = ggplot(alphas,aes(alpha, fit_alpha)) +
  geom_point(color = 'grey33', size = 5) +
  geom_smooth(method='lm', se=FALSE, color='grey33') +
  theme_few() +
  labs(x='Simulated Alpha', y='Fit Alpha', title='') +
  theme(plot.title = element_text(hjust=0.5, size=20, face='bold')) + 
  theme(axis.text = element_text(size = 16))


```


```{r alpha comparison across model 1 & model 2}

lr_models <- read.csv('../RP_Learning/models/lr_both_models.csv')

ggscatter(lr_models, x = "fit_alpha", y = "fit_punishment_alpha", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Model 1 Learning Rate", ylab = "Model 2 Punishment Learning Rate")

# is the data normally distributed-- no, the p values are less than 0.05 -- switch to non-parametric test

shapiro.test(lr_models$fit_alpha) # => p = 0.00061
# Shapiro-Wilk normality test for wt
shapiro.test(lr_models$fit_punishment_alpha) # => p = 5.122e-05

# test: S = 39106, p < 2.2e-16

res <- cor.test(lr_models$fit_alpha, lr_models$fit_punishment_alpha, 
                    method = "spearman")

res$p.value # 0.017

res


### plot 

library(ggplot2)
library(ggthemes)

plot = ggplot(lr_models,aes(fit_alpha, fit_punishment_alpha)) +
  geom_point(color = 'grey33', size = 5) +
  geom_smooth(method='lm', se=FALSE, color='grey33') +
  theme_few() +
  labs(x='Model 1 Learning Rate', y='Model 2 Punishment Learning Rate', title='') +
  theme(plot.title = element_text(hjust=0.5, size=20, face='bold')) + 
  theme(axis.text = element_text(size = 16))

```

# Model 1 Group differences 

no significant differences between learning rates (effect size: 0.0795, p=0.56) 
```{r}

# Summary statistics
data_m1 %>%
  group_by(group) %>%
  get_summary_stats(value, type = "mean_sd")

# test for normalcy: doesn't meet criteria

with(data_m1, shapiro.test(value[group == "HC"])) # p = 5.35 e -0.5
with(data_m1, shapiro.test(value[group == "MDD"])) # p = 6.85 e-06

# variances: F=1.579 p = 0.166 so no significant difference between the variances 
res.ftest <- var.test(value ~ group, data = data_m1)
res.ftest

res <- t.test(value ~ group, data = data_m1) # t = 1.617 p= 0.110
res

# using an unpaired two-samples Wilcoxon test, we conclude there is no significant difference in learning rates between HC and MDD (W=913.5, p=0.0797)
wilcox.test(value ~ group, data = data_m1)

# wilcox test
stat.test <- data_m1 %>% 
  wilcox_test(value ~ group) %>%
  add_significance()
stat.test

# calculate effect size: small -- 0.200
data_m1 %>% wilcox_effsize(value ~ group)

```

```{r plotting group differences}

# summary stats
m1 <- data_summary(data_m1, varname="value", 
                    groupnames=c("group"))

p <- ggplot(m1, aes(x=group, y=value, fill=group)) + 
   geom_bar(stat="identity", position=position_dodge()) +
  geom_errorbar(aes(ymin=value-sd, ymax=value+sd), width=.2,
                 position=position_dodge(.9))
  
p + scale_fill_brewer(palette="Paired") + theme_minimal()
```

```{r beta}

# Summary statistics
m1_beta %>%
  group_by(group) %>%
  get_summary_stats(fit_beta, type = "mean_sd")

# test for normalcy: doesn't meet criteria

with(m1_beta, shapiro.test(fit_beta[group == "HC"])) # p = 1.14e-07
with(m1_beta, shapiro.test(fit_beta[group == "MDD"])) # p = 4.61e-08

# variances: F=1.083 p = 0.808so no significant difference between the variances 
res.ftest <- var.test(fit_beta ~ group, data = m1_beta)
res.ftest

res <- t.test(fit_beta ~ group, data = m1_beta) # t = 0.548 p= 0.585
res

# using an unpaired two-samples Wilcoxon test, we conclude there is no significant difference in learning rates between HC and MDD (W=807.5, p=0.495)
wilcox.test(fit_beta ~ group, data = m1_beta)

# wilcox test
stat.test <- m1_beta %>% 
  wilcox_test(fit_beta ~ group) %>%
  add_significance()
stat.test

# calculate effect size
m1_beta %>% wilcox_effsize(fit_beta ~ group) # small effect = 0.0784, p = 0.495

```

```{r plotting beta}
# summary stats
m1_beta_sum <- data_summary(m1_beta, varname="fit_beta", 
                    groupnames=c("group"))

p <- ggplot(m1_beta_sum, aes(x=group, y=fit_beta, fill=group)) + 
   geom_bar(stat="identity", position=position_dodge()) +
  geom_errorbar(aes(ymin=fit_beta-sd, ymax=fit_beta+sd), width=.2,
                 position=position_dodge(.9))
  
p + scale_fill_brewer(palette="Paired") + theme_minimal()

```


# Model 2 Group differences

no significant group differences in reward LR -- small effect size(0.0852), p = 0.533

```{r reward LR}

reward_alpha = data_m2 %>% filter(parameter == "fit_reward_alpha")
p_alpha = data_m2 %>% filter(parameter == "fit_punishment_alpha")

# Summary statistics
reward_alpha %>%
  group_by(group) %>%
  get_summary_stats(value, type = "mean_sd")

# test for normalcy: doesn't meet criteria

with(reward_alpha, shapiro.test(value[group == "HC"]))# p = 0.0043
with(reward_alpha, shapiro.test(value[group == "MDD"])) # p = 0.0047

# variances: F=0.99 p = 0.98 so no significant difference between the variances 
res.ftest <- var.test(value ~ group, data = reward_alpha)
res.ftest

res <- t.test(value ~ group, data = reward_alpha) # t = 1.01 p= 0.315
res

# using an unpaired two-samples Wilcoxon test, we conclude there is no significant difference in learning rates between HC and MDD (W=862, p=0.219)
wilcox.test(value ~ group, data = reward_alpha)

# wilcox test
stat.test <- reward_alpha %>% 
  wilcox_test(value ~ group) %>%
  add_significance()
stat.test

# calculate effect size
reward_alpha %>% wilcox_effsize(value ~ group) # effect size = 0.141, p = 0.219

```

no significant difference: small effect size (0.00911) p=0.953
```{r punishment LR}

# Summary statistics
p_alpha %>%
  group_by(group) %>%
  get_summary_stats(value, type = "mean_sd")

# test for normalcy: doesn't meet criteria

with(p_alpha, shapiro.test(value[group == "HC"]))# p = 1.498e-07
with(p_alpha, shapiro.test(value[group == "MDD"])) # p = 3.027e-08

# variances: F=1.706 p = 0.106 so no significant difference between the variances 
res.ftest <- var.test(value ~ group, data = p_alpha)
res.ftest

res <- t.test(value ~ group, data = p_alpha) # t = 0.995 p= 0.323
res

# using an unpaired two-samples Wilcoxon test, we conclude there is no significant difference in learning rates between HC and MDD (W=809.5, p=0.488)
wilcox.test(value ~ group, data = p_alpha)

# wilcox test
stat.test <- p_alpha %>% 
  wilcox_test(value ~ group) %>%
  add_significance()
stat.test

# calculate effect size
p_alpha %>% wilcox_effsize(value ~ group) # effect size = 0.0796 p=0.488
```

LR across reward and punishment show a moderate effect size of 0.343, p = 0.0028
```{r HC reward vs punishment}

m2_hc <- lrs %>% filter(group == "HC")


m2_hc_m <-
  m2_hc %>% select("subject_ID",
                   "group",
                   "fit_reward_alpha",
                   "fit_punishment_alpha") %>% reshape2::melt(
                          id.vars = c("subject_ID", "group"),
                          variable.name = "parameter")
# Summary statistics
m2_hc_m %>%
  group_by(parameter) %>%
  get_summary_stats(value, type = "mean_sd")

# test for normalcy: doesn't meet criteria

with(m2_hc_m, shapiro.test(value[parameter == "fit_reward_alpha"])) # p = 0.0047
with(m2_hc_m, shapiro.test(value[parameter == "fit_punishment_alpha"])) # p = 3.03e-07

# variances: F=1.1007 p = 0.772 so no significant difference between the variances 
res.ftest <- var.test(value ~ parameter, data = m2_hc_m)
res.ftest

res <- t.test(value ~ parameter, data = m2_hc_m) # t = 2.71 p= 0.008
res

# using an unpaired two-samples Wilcoxon test, we conclude there is no significant difference in learning rates between HC and MDD (W=1010, p=0.0028)
wilcox.test(value ~ parameter, data = m2_hc_m)

# wilcox test
stat.test <- m2_hc_m %>% 
  wilcox_test(value ~ parameter) %>%
  add_significance()
stat.test

# calculate effect size
m2_hc_m %>% wilcox_effsize(value ~ parameter) # moderate effect size = 0.343

```


moderate effect size (0.310) p = 0.0209
```{r MDD RvP differences}

m2_mdd <- lrs%>% filter(group == "MDD")

m2_mdd_m <-
  m2_mdd %>% select("subject_ID",
                   "group",
                   "fit_reward_alpha",
                   "fit_punishment_alpha") %>% reshape2::melt(
                          id.vars = c("subject_ID", "group"),
                          variable.name = "parameter")


# Summary statistics
m2_mdd_m %>%
  group_by(parameter) %>%
  get_summary_stats(value, type = "mean_sd")

# test for normalcy: doesn't meet criteria

with(m2_mdd_m, shapiro.test(value[parameter == "fit_reward_alpha"])) # p = 0.00093
with(m2_mdd_m, shapiro.test(value[parameter == "fit_punishment_alpha"])) # p = 5.2e-08

# variances: F=1.95 p = 0.042, significant difference between the variances 
res.ftest <- var.test(value ~ parameter, data = m2_mdd_m)
res.ftest

res <- t.test(value ~ parameter, data = m2_mdd_m) # t = 3.23 p= 0.0019
res

wilcox.test(value ~ parameter, data = m2_mdd_m)

# wilcox test, W = 940, p =0.02
stat.test <- m2_mdd_m %>% 
  wilcox_test(value ~ parameter) %>%
  add_significance()
stat.test

# calculate effect size
m2_mdd_m %>% wilcox_effsize(value ~ parameter) # small effect size = 0.259
```

```{r plotting LR results}

# summary stats
hc_lrs <- m2_hc_m %>%
  group_by(parameter) %>%
  get_summary_stats(value, type = "mean_sd")

mdd_lrs <- m2_mdd_m %>%
  group_by(parameter) %>%
  get_summary_stats(value, type = "mean_sd")

# add group column 
hc_lrs$group <- "HC"
mdd_lrs$group <- "MDD"

# rbind 

group_lrs <- rbind(hc_lrs, mdd_lrs)

# plot

p <- ggplot(group_lrs, aes(x = parameter, y = mean, fill = group)) +
  labs(x = "",
       y = "mean learning rate",
       title = "") +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd),
                width = .2,
                position = position_dodge(.9)) +
  scale_fill_manual(values=c('lightblue3','peachpuff')) +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 20),
    axis.title = element_text(size = 24),
    legend.title = element_blank(),
    legend.text = element_text(size=20),
    legend.position = c(0.90, 0.95),
  )


```

```{r plotting punishment LR}

# summary stats
punishment <- data_summary(p_alpha, varname="value", 
                    groupnames=c("group"))


p <- ggplot(punishment, aes(x=group, y=value, fill=group)) + 
   geom_bar(stat="identity", position=position_dodge()) +
  geom_errorbar(aes(ymin=value-sd, ymax=value+sd), width=.2,
                 position=position_dodge(.9))
  
p + scale_fill_brewer(palette="Paired") + theme_minimal()


```

```{r beta}

# Summary statistics
m2_beta %>%
  group_by(group) %>%
  get_summary_stats(fit_beta, type = "mean_sd")

# test for normalcy: doesn't meet criteria

with(m2_beta, shapiro.test(fit_beta[group == "HC"])) # p = 7.197e-07
with(m2_beta, shapiro.test(fit_beta[group == "MDD"])) # p = 8.53e-07

# variances: F=1.057 p = 0.864 so no significant difference between the variances 
res.ftest <- var.test(fit_beta ~ group, data = m2_beta)
res.ftest

res <- t.test(fit_beta ~ group, data = m2_beta) # t = 0.394 p= 0.695
res

# using an unpaired two-samples Wilcoxon test, we conclude there is no significant difference in learning rates between HC and MDD (W=806, p=0.507
wilcox.test(fit_beta ~ group, data = m2_beta)

# wilcox test
stat.test <- m2_beta %>% 
  wilcox_test(fit_beta ~ group) %>%
  add_significance()
stat.test

# calculate effect size
m2_beta %>% wilcox_effsize(fit_beta ~ group) # small effect size = 0.076, p =0.507
```

```{r}

# summary stats
m2_beta_sum <- data_summary(m2_beta, varname="fit_beta", 
                    groupnames=c("group"))

p <- ggplot(m2_beta_sum, aes(x=group, y=fit_beta, fill=group)) + 
   geom_bar(stat="identity", position=position_dodge()) +
  geom_errorbar(aes(ymin=fit_beta-sd, ymax=fit_beta+sd), width=.2,
                 position=position_dodge(.9))
  
p + scale_fill_brewer(palette="Paired") + theme_minimal()


```

